### 2018-10-29 Paper/Paper_BillietStijn_RadeonRays.tex
Added paper to repository seeing I might need to revert from time to time :D.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% LaTeX book template                           %%
%% Author:  Amber Jain (http://amberj.devio.us/) %%
%% License: ISC license                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[a4paper,12pt,oneside]{book}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Source: http://en.wikibooks.org/wiki/LaTeX/Hyperlinks %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{titlepic}
\usepackage{wrapfig}

\graphicspath{{./images/}}
\setlength\parindent{0pt}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% First page of book which contains 'stuff' like: %
%  - Book title, subtitle                         %
%  - Book author name                             %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Book's title and subtitle
\title{
	\Huge \textbf{Hybrid rendering}\\
	\huge using Radeon Rays
}
\titlepic{\includegraphics[width=\textwidth]{radeonrays.jpg}}
% Author
\author{
	Stijn Billiet\\\\
	Graduation work 2018-19\\
	Digital Arts And Entertainment\\
	howest.be
}
\date{}

\begin{document}

\frontmatter
\maketitle
\cleardoublepage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Auto-generated table of contents, list of figures and list of tables %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\tableofcontents

\mainmatter
\chapter{Abstract}
This paper was written in an effort to elaborate on hybrid solutions between modern day rasterization engines and powerful ray-tracing ones and to explain rather thoroughly how one could implement this.

TODO methods
TODO results

\chapter{Introduction}
Ray-tracing is usually used in VFX productions to create the most accurate lighting effects, this doesn’t come for free though seeing crisp ray-traced images require countless hours of rendering. With the rise of new ray-tracing capable hardware (e.g. Volta architecture NVIDIA) things are about to change, developers are already putting ray-tracing in some parts of their pipelines, they must however rely on optimization (e.g. De-noising) to keep up with the performance demands. So for the time being we will see a lot of hybrid solutions, but we should see full ray traced engines by the time that our hardware is capable enough.\newline\par

That being said I would like to clarify that rasterization engines aren’t bad, however they are very rough approximations and rely on specific techniques to achieve lighting effects sometimes at the cost of artifacts. We must keep in mind though rasterization engines are here to stay for a while longer due to their immense performance, and not to forget they have been around for a while, their maturity surely pays off seeing what we’ve already achieved with rasterization (e.g. The Witcher 3: Wild Hunt).\newline\par

As the subtitle suggests we will use Radeon Rays for our ray-tracing algorithm, however there are other options out there like Microsoft's DirectX Raytracing. Nevertheless the topics outlined should still hold up when implementing an other API, only the Radeon Rays implementation details would be an exception to that rule. The same essentialy goes for our graphics card specification, OpenGL was used seeing that would be the logical combination, but the information should translate over nicely to other API's. \newline\par

%%%%%%%%%%%%%%%%
% NEW CHAPTER! %
%%%%%%%%%%%%%%%%
\chapter{Research}

This extract should give you a rather thorough overview of the technologies used, however some knowledge regarding graphics programming and/or matrix mathematics is advised.\newline\par

\section{Rendering techniques}
Over the past century many rendering algorithms have been researched/employed. They are all trying to find an answer to the problem of limited rendering capabilities. Seeing that our modern model of light would be completely impractical when it comes to (high speed) simulations, meaning it would take an obscene amount of time to calculate.\newline\par

Therefore people have come up with ways to more efficiently model the phenomenon of light. The most prominent models being: ray tracing (ray casting) and rasterization. The aforementioned models will be covered in the following subsections.

\pagebreak

\subsection{Raytracing}
\includegraphics[width=\textwidth]{raytracingtest.png}
\subsubsection{Definition}
\begin{minipage}{0.3\textwidth}
	\includegraphics[width=\textwidth]{raytracing.png}
\end{minipage}
\begin{minipage}{0.7\textwidth}
	In short ray-tracing is a rendering technique that can produce lifelike lighting effects. A ray-tracing algorithm essentially achieves this by sending out rays for every pixel of our framebuffer (camera) towards the corresponding scene geometry. Based on the technique we can send out additional bounces (secondary rays). It then basically backtracks that chain of light-rays to accumulate material/light information to then calculate the final pixel color.
\end{minipage}

\subsubsection{Ray casting}

As addressed before ray-tracing essentially probes material and light information through the use of rays. There are 2 sets of rays, we've got primary rays and secondary rays. Primary rays are the first rays being sent out, the rays from camera towards the scene geometry. These primary rays are essentially doing the same as with rasterization in the sense that they only probe at material information. It becomes interesting however when we cast our secondary rays, these secondary rays have their origins at the interception point of the primary rays and the scene geometry. Their direction on the other hand are totally dependent on the technique that is employed, meaning that we will use different techniques to bounce/direct them based on the effect required  (e.g. shadows, reflections, etc.).\newline\par

\pagebreak

\subsection{Rasterization}

\begin{figure}[h]
\includegraphics[width=\textwidth]{rasterization.jpg}
\caption{Rasterization graphics pipeline.}
\end{figure}

\subsubsection{Definition}
Rasterization is an algorithm designed to display three-dimensional objects on a two-dimensional screen. It's very fast seeing this algorithm is parallel in nature and can thus make full use of the GPU. The results of rasterization have gotten very good over the years, very convincing at relatively low compute performance. However there are certain limitations to this algorithm, seeing it's main goal is to get graphics in screen space.\newline\par

\subsubsection{Graphics pipeline}
With rasterization objects on screen are created from meshes. A mesh is a collection of three-dimensional points a.k.a. vertices. These vertices are the main resource needed for rendering, these go through what's called the rendering pipeline to then be converted into screen pixels. The rendering pipeline or more formerly known as the graphics pipeline acts as a conceptual model that describes what steps a graphics system needs to performs to be able to render aforementioned meshes to a 2D screen. \newline\par
\pagebreak

\subsubsection{Vertex shader}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{coordinate_systems.png}
	\caption{Model space to screen space.}
\end{figure}
The first stage is the vertex shader which essentially transforms the vertices to their correct
positions, the vertices have to be transformed seeing that they are defined in their own local transform space(vertices positions are defined relative to their own center point). The vertices are defined this way as a means to make it easier for modelers, they quite simply have to define the vertices relative to the modeling software's gizmo.

\subsubsection{Shape assembly}
\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{primitives.png}
	\caption{List of possible OpenGL primitives.}
\end{figure}

The second stage is the shape assembly which is in charge of creating primitives out of these verts. The developers have to specify the wanted primitive (e.g. triangle) the GPU then uses that information to interpret the a collection of vertices as said primitive.\newline\par

\pagebreak

\subsubsection{Geometry shader}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.7\textwidth]{geometryshader.jpg}
	\caption{Purpose of geometry shader.}
\end{figure}
The third stage is an optional geometry shader, this stage won't be covered in depth, but it is essentially in charge of additional geometry creation (extra vertices).\newline\par

\subsubsection{Rasterization stage}
The fourth stage is rasterization, which interpolates between vertex information (e.g. vertex color) and subsequently rasterizing the image into pixel fragments. \newline\par

*culling
*clipping
*homogenous divide
*viewport transform
*rasterization

\subsubsection{Fragment shader}
The fifth stage then applies shading to those pixel fragments.

\subsubsection{Test and blending stage}
Finally we've got the testing and blending stage.

\pagebreak

\section{Hybrid rendering approach}
\begin{figure}[h]
	\includegraphics[width=\textwidth]{hybrid.jpg}
	\caption{Hybrid rendering empowered by Nvidia RTX}
\end{figure}
The different rendering techniques have been covered, now onto a hybrid solution where rasterization and raytracing are combined to achieve a new pipeline. Most hybrid approaches replace the first stage of ray-tracing, which is sending out our primary rays to retrieve surface information, by rasterization. This effectively reduces the amount of rays by half. Ray-tracing comes into play for sending out our reflected rays, which could be anything from sending out new primary rays to secondary (shadow) rays. This introduction of raytracing essentially brings us world information whereas normally you'd only have surface information. To be able to augment rasterization with raytracing we will need to postpone our lighting effects to a later stage. In a forward rendering pipeline the lighting is done per mesh in the fragment shader stage, effectively disabling us from hooking raytracing to that pipeline.

\pagebreak
\subsection{Raytracing in rasterization}

\subsubsection{Deferred shading}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Duis risus ante, auctor et pulvinar non, posuere ac lacus. Praesent egestas nisi id metus rhoncus ac lobortis sem hendrerit. Etiam et sapien eget lectus interdum posuere sit amet ac urna. Aliquam pellentesque imperdiet erat, eget consectetur felis malesuada quis. Pellentesque sollicitudin, odio sed dapibus eleifend, magna sem luctus turpis, id aliquam felis dolor eu diam. Etiam ullamcorper, nunc a accumsan adipiscing, turpis odio bibendum erat, id convallis magna eros nec metus. Sed vel ligula justo, sit amet vestibulum dolor. Sed vitae augue sit amet magna ullamcorper suscipit. Quisque dictum ipsum a sapien egestas facilisis.

\subsubsection{G-Buffers}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Duis risus ante, auctor et pulvinar non, posuere ac lacus. Praesent egestas nisi id metus rhoncus ac lobortis sem hendrerit. Etiam et sapien eget lectus interdum posuere sit amet ac urna. Aliquam pellentesque imperdiet erat, eget consectetur felis malesuada quis. Pellentesque sollicitudin, odio sed dapibus eleifend, magna sem luctus turpis, id aliquam felis dolor eu diam. Etiam ullamcorper, nunc a accumsan adipiscing, turpis odio bibendum erat, id convallis magna eros nec metus. Sed vel ligula justo, sit amet vestibulum dolor. Sed vitae augue sit amet magna ullamcorper suscipit. Quisque dictum ipsum a sapien egestas facilisis.

\subsection{Raytraced shadows}
\begin{wrapfigure}{r}{0.4\textwidth}
	\includegraphics[width=0.4\textwidth]{shadows.jpg}
\end{wrapfigure}
For example when looking at ray-traced shadows. To determine if a pixel should be shaded or not a primary ray is cast from that pixel into the scene, material information is then gathered and additional secondary rays are being cast towards the relevant lights. If that secondary ray/shadow ray is occluded by another object then that pixel will be shaded using only the material information, if it isn't occluded however we can add the target lights information to to the shaders equation.\newline\par

\subsubsection{Soft shadows}
Test test test

\subsection{Optimizations}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Duis risus ante, auctor et pulvinar non, posuere ac lacus. Praesent egestas nisi id metus rhoncus ac lobortis sem hendrerit. Etiam et sapien eget lectus interdum posuere sit amet ac urna. Aliquam pellentesque imperdiet erat, eget consectetur felis malesuada quis. Pellentesque sollicitudin, odio sed dapibus eleifend, magna sem luctus turpis, id aliquam felis dolor eu diam.

%%%%%%%%%%%%%%%%
% NEW CHAPTER! %
%%%%%%%%%%%%%%%%
\chapter{Case study}
Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.

\section{Implementation rasterization}
\subsection{Window creation}
\subsection{Rendering specification}
\subsection{Shader loading}
\subsection{Coordinate systems}
\subsection{Model loading}
\subsection{Texture loading}
\subsection{Deferred shading}
\section{Implementation ray-tracing}
\subsection{Radeon rays implementation}
\subsection{Shadow implementation}
\subsection{Optimizations}

\chapter{Conlusion}
Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.

\bibliographystyle{}
\bibliography{}
\begin{thebibliography}{Xyz12}

	\bibitem{voica14}
	Alexandru Voica,
	2014,
	Practical techniques for ray tracing in games,
	\url{http://www.gamasutra.com/blogs/AlexandruVoica/20140318/213148/Practical_techniques_for_ray_tracing_in_games.php}

	\bibitem{voica15}
	Alexandru Voica,
	2015,
	Ray tracing made easy,
	\url{http://www.alexvoica.com/ray-tracing-made-easy/#sthash.q1sHqqiT.dpbs}

	\bibitem{einig17}
	Mathieu Einig,
	2017,
	Hybrid rendering for real-time lighting: ray tracing vs rasterization,
	\url{https://www.imgtec.com/blog/hybrid-rendering-for-real-time-lighting/}

	\bibitem{caulfield18}
	Brian Caulfield,
	2018,
	What's the difference between ray tracing and rasterization?,
	\url{https://blogs.nvidia.com/blog/2018/03/19/whats-difference-between-ray-tracing-rasterization/}

	\bibitem{pipeline18}
	Wikipedia,
	2018,
	Graphics pipeline,
	\url{https://en.wikipedia.org/wiki/Graphics_pipeline}

	\bibitem{scratchapixel18}
	Scratchapixel,
	2018,
	Ray-Tracing: Generating Camera Rays,
	\url{https://www.scratchapixel.com/lessons/3d-basic-rendering/ray-tracing-generating-camera-rays}
\end{thebibliography}

%%%%%%%%%%%%%%%%
% NEW CHAPTER! %
%%%%%%%%%%%%%%%%
\chapter*{Appendices}

\end{document}

### 2018-10-29 Paper/Paper_BillietStijn_RadeonRays.tex
Fixed minor issue with duplicate picture and removed minipage near raytracing part

@@
 \subsection{Raytracing}
-\includegraphics[width=\textwidth]{raytracingtest.png}
+
+\begin{figure}[h]
+	\caption{Illustration of different ray-casting solutions}
+	\centering
+	\includegraphics[width=0.95\textwidth]{raycasting.jpg}
+\end{figure}
 \subsubsection{Definition}
-\begin{minipage}{0.3\textwidth}
-	\includegraphics[width=\textwidth]{raytracing.png}
-\end{minipage}
-\begin{minipage}{0.7\textwidth}
 In short ray-tracing is a rendering technique that can produce lifelike lighting effects.
@@
 It then basically backtracks that chain of light-rays to accumulate material/light information to then calculate the final pixel color.
-\end{minipage}
 
@@
 \end{figure}
-The third stage is an optional geometry shader, this stage won't be covered in depth, but it is essentially in charge of additional geometry creation (extra vertices).\newline\par
+The third stage is an optional geometry shader, this stage won't be covered in depth, but it is essentially in charge of creating additional geometry (extra vertices) on the fly (at runtime).\newline\par
 
 \subsubsection{Rasterization stage}
-The fourth stage is rasterization, which interpolates between vertex information (e.g. vertex color) and subsequently rasterizing the image into pixel fragments. \newline\par
+The fourth stage is rasterization, this is where the algorithm gets it's name.\newline\par
 

### 2018-10-29 Paper/Paper_BillietStijn_RadeonRays.tex
Worked on part about homogeneous divide and rasterization

@@
 	\centering
-	\includegraphics[width=0.7\textwidth]{geometryshader.jpg}
+	\includegraphics[width=0.5\textwidth]{geometryshader.jpg}
 	\caption{Purpose of geometry shader.}
@@
 \subsubsection{Rasterization stage}
-The fourth stage is rasterization, this is where the algorithm gets it's name.\newline\par
+The fourth stage is rasterization, this is where the algorithm gets it's name.
+The rasterizer is in charge of a not only rasterization, but also homogeneous divide, clipping, culling and viewport transform.
+\newline\par
+
+\begin{minipage}{0.6\textwidth}
+	The first task of the rasterizer is the homogeneous divide.
+The previous stages were in charge of transforming the vertices from model space into clipping space, this is done by multiplying the world, view and projection matrix with the vertices this yields a valid projection but does not account for depth information (f.e.
+objects that are nearby will appear the same as objects from afar) this is where the z-component comes in, seeing the z stores our depth information.
+Before we apply the projectionmatrix we store the z-component as an additional w component.
+\end{minipage} \hfill
+\begin{minipage}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{frustrum.png}
+\end{minipage}
+\vspace{0.2cm}
+
+The homogeneous divide is just a specific term for converting the homogeneous coordinate back to w=1 meaning we put the vertices in their normalized device coordinates (in the -1 to 1 range) effectively scaling all objects to their rightful dimensions.
+
+\begin{minipage}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{pixelisation.png}
+\end{minipage} \hfill
+\begin{minipage}{0.55\textwidth}
+	Rasterization(pixelisation) is the process of converting continuous primitives into discrete pixel fragments.
+The rasterizer does so by interpolating between vertex attributes (data that vertex holds) like color, normals, texture coordinates and positions.
+It then stores that information in the pixel fragments before finally sending it off to the pixel/fragment shader to be shaded.
+\end{minipage}
+\vspace{0.5cm}
 

### 2018-10-29 Paper/Paper_BillietStijn_RadeonRays.tex
Altered order of certain stages to improve readability

@@
 
-\begin{minipage}{0.6\textwidth}
+\begin{minipage}{0.59\textwidth}
 	The first task of the rasterizer is the homogeneous divide.
@@
 \end{minipage}
-\vspace{0.2cm}
+\vspace{0.25cm}
 
@@
 
+\pagebreak
+
 \begin{minipage}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{pixelisation.png}
+	\includegraphics[width=0.8\textwidth]{culling.png}
 \end{minipage} \hfill
+\begin{minipage}{0.59\textwidth}
+	The second task is the clipping stage, in this stage we clip triangles that lie (partly) outside of the view frustum.
+This effectively removes those triangles and prevents them from being projected/rendered.
+\end{minipage}
+
+\vspace{0.5cm}
+
 \begin{minipage}{0.55\textwidth}
@@
 It then stores that information in the pixel fragments before finally sending it off to the pixel/fragment shader to be shaded.
+\end{minipage} \hfill
+\begin{minipage}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{pixelisation.png}
 \end{minipage}

### 2018-10-30 Paper/Paper_BillietStijn_RadeonRays.tex
Finished part about rasterization

--- 2018-10-29 Paper/Paper_BillietStijn_RadeonRays.tex
@@
 	\caption{Model space to screen space.}
+	\label{fig:coordinates}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=0.5\textwidth]{geometryshader.jpg}
+	\includegraphics[width=0.7\textwidth]{geometryshader.jpg}
 	\caption{Purpose of geometry shader.}
@@
 
-\begin{minipage}{0.6\textwidth}
-	The first task of the rasterizer is the homogeneous divide.
+\begin{minipage}{0.5\textwidth}
+	\centering
+	\includegraphics[width=\textwidth]{culling.png}
+\end{minipage} \hfill
+\begin{minipage}{0.5\textwidth}
+	The first task of the rasterizer is primitive culling.
+The shapes that made it to the rasterizer are checked based on their normals.
+The normal of a primitive/vertex tells us if the face is facing away from us or not.
+This simple operation can be demonstrated using the dotproduct, however this does get handled by the specifications (e.g.
+DirectX/OpenGL) rasterizer.
+In it's essence the calculation will translate to if the dotproduct is negative then the face is facing away, if it's positive then it's facing toward us.
+\end{minipage}
+
+\pagebreak
+
+\begin{minipage}{0.5\textwidth}
+	The second task is the clipping stage, in this stage we clip triangles that lie (partly) outside of the view frustum.
+This effectively removes those triangles and prevents them from being projected/rendered.
+\end{minipage} \hfill
+\begin{minipage}{0.5\textwidth}
+	\centering
+	\includegraphics[width=0.8\textwidth]{clipping.png}
+\end{minipage}
+
+\vspace{0.5cm}
+
+\begin{minipage}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{frustrum.png}
+\end{minipage} \hfill
+\begin{minipage}{0.59\textwidth}
+	The third task of the rasterizer is the homogeneous divide.
 The previous stages were in charge of transforming the vertices from model space into clipping space, this is done by multiplying the world, view and projection matrix with the vertices this yields a valid projection but does not account for depth information (f.e.
@@
 Before we apply the projectionmatrix we store the z-component as an additional w component.
-\end{minipage} \hfill
-\begin{minipage}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{frustrum.png}
 \end{minipage}
-\vspace{0.2cm}
 
-The homogeneous divide is just a specific term for converting the homogeneous coordinate back to w=1 meaning we put the vertices in their normalized device coordinates (in the -1 to 1 range) effectively scaling all objects to their rightful dimensions.
+\vspace{0.25cm}
+
+The homogeneous divide is just a specific term for converting the homogeneous coordinate back to w=1 meaning we put the vertices in their normalized device coordinates (in the -1 to 1 range) effectively scaling all objects to their rightful dimensions.\newline\par
+
+The fourth stage is the viewport transform.
+In the viewport transform we essentially translate all the vertex positions that are now in normalized device coordinates (NDC) and transform them to window space.
+These are then the coordinates that are being rasterized to the output image.
+This process is outlined in figure \ref{fig:coordinates} (part with green outline).
+
+\vspace{0.5cm}
 
 \begin{minipage}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{pixelisation.png}
+	\includegraphics[width=0.95\textwidth]{pixelisation.png}
 \end{minipage} \hfill
-\begin{minipage}{0.55\textwidth}
-	Rasterization(pixelisation) is the process of converting continuous primitives into discrete pixel fragments.
+\begin{minipage}{0.59\textwidth}
+	The final task pixelisation (rasterisation) is the process of converting continuous primitives into discrete pixel fragments.
 The rasterizer does so by interpolating between vertex attributes (data that vertex holds) like color, normals, texture coordinates and positions.
@@
 \end{minipage}
-\vspace{0.5cm}
 
-*culling
-*clipping
-*homogenous divide
-*viewport transform
-*rasterization
+\pagebreak
 
@@
 \subsection{Raytracing in rasterization}
+
+\subsubsection{Forward vs deferred rendering}
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Duis risus ante, auctor et pulvinar non, posuere ac lacus.
+Praesent egestas nisi id metus rhoncus ac lobortis sem hendrerit.
+Etiam et sapien eget lectus interdum posuere sit amet ac urna.
+Aliquam pellentesque imperdiet erat, eget consectetur felis malesuada quis.
+Pellentesque sollicitudin, odio sed dapibus eleifend, magna sem luctus turpis, id aliquam felis dolor eu diam.
+Etiam ullamcorper, nunc a accumsan adipiscing, turpis odio bibendum erat, id convallis magna eros nec metus.
+Sed vel ligula justo, sit amet vestibulum dolor.
+Sed vitae augue sit amet magna ullamcorper suscipit.
+Quisque dictum ipsum a sapien egestas facilisis.
 

### 2018-11-02 Paper/Paper_BillietStijn_RadeonRays.tex
Worked on research part of paper, almost done then onto case study

@@
 \end{figure}
-The third stage is an optional geometry shader, this stage won't be covered in depth, but it is essentially in charge of creating additional geometry (extra vertices) on the fly (at runtime).\newline\par
+The third stage is an optional geometry shader, this stage won't be covered in depth, but it is essentially in charge of creating additional geometry (extra vertices) at runtime.\newline\par
 
@@
 
+\begin{minipage}{0.55\textwidth}
+	The final task pixelisation (rasterisation) is the process of converting continuous primitives into discrete pixel fragments.
+The rasterizer does so by interpolating between vertex attributes (data that vertex holds) like color, normals, texture coordinates and positions.
+It then launches a fragment shader for every pixel fragment with the information just calculated.
+\end{minipage} \hfill
 \begin{minipage}{0.4\textwidth}
 	\includegraphics[width=0.95\textwidth]{pixelisation.png}
-\end{minipage} \hfill
-\begin{minipage}{0.59\textwidth}
-	The final task pixelisation (rasterisation) is the process of converting continuous primitives into discrete pixel fragments.
-The rasterizer does so by interpolating between vertex attributes (data that vertex holds) like color, normals, texture coordinates and positions.
-It then stores that information in the pixel fragments before finally sending it off to the pixel/fragment shader to be shaded.
 \end{minipage}
@@
 \subsubsection{Fragment shader}
-The fifth stage then applies shading to those pixel fragments.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.85\textwidth]{Zbuffer.png}
+\end{figure}
+The fifth stage, the fragment stage is a shader stage that will process all pixel fragments processed by the rasterizer.
+Depending on the technique we collapse all information into a set of colors (color buffer) and a single depth value (depth buffer), additionally we can augment this by writing away a stencil value (stencil buffer).
 
 \subsubsection{Test and blending stage}
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.8\textwidth]{alpha.png}
+\end{figure}
 Finally we've got the testing and blending stage.
+Up until this point "pixels" where addressed as pixel fragments, this is in-fact a clear distinction to make seeing that pixel fragments are not guaranteed to show up on screen, they could be discarded in the shaders or they could be transparent.
+If there are transparent objects in the scene we have to keep both pixels, those from the object behind it and the transparent object, we can then simply blend those pixels together to receive the final pixel colour.
+This blending operation is in-fact quite costly seeing we need to make sure the objects appear in the right order and we've got the load of additional pixels on our hands.
+A more optimized technique can be used for objects that are fully transparent however, here we can employ alpha testing where we simply discard those pixel fragments in the fragment shader.
 
@@
 
+\section{API's and tools}
+
+The main API's that have been researched were Radeon Rays, OpenCL and OpenGL.
+However to speed up the development process tools were employed.
+The main part of the tools were tailored towards the OpenGL pipeline (e.g.
+model loading, texture loading, etc.), but some small frameworks were also employed to speed up the raytracer development (e.g.
+CLW).
+Keep in mind however that these tools are not necessary seeing it could all be made on your own, but it sure helps to have these around.
+
+\subsection{Raytracing}
+\subsubsection{OpenCL}
+Open Computing Language is a framework for writing programs that execute across compute devices.
+This includes among other hardware accelerators, the central processing unit (CPU) and the graphics processing unit (GPU).
+These aforementioned programs are called kernels and are also written on a pseudo c-language.
+OpenCL provides the programmer of a standard interface to perform parallel computing tasks.
+The programmer has to set up the OpenCL execution pipeline and write the kernels.
+Under this section you'll find a quick rundown of aforementioned execution
+model.\newline
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.9\textwidth]{openclprogram.png}
+\end{figure}
+
+\pagebreak
+
+\begin{itemize}
+	\item{Platform}\newline
+	The platform that we are running this code on, so which vendor (e.g. Nvidia).
+	\item{Device}\newline
+	What device are we going to run this on could be anything from CPU, GPU to APU.
+	\item{Context}\newline
+	This is the runtime interface between the host program (run on the CPU) and the device (CPU/GPU/etc.).
+The context is in charge of managing all the OpenCL resources so programs, kernels, command queues and buffers.
+	\item{Program}\newline
+	The entire OpenCL program, this could be one or more kernels/device functions bundled together into a single program file.
+	\item{Kernel}\newline
+	The starting point/building blocks of a OpenCL program.
+Kernels are called from the CPU, they provide the basic units of executable code that run on a OpenCL device.
+The kernels are preceded with the keyword "\_\_kernel"
+	\item{Command queue}\newline
+	The command queue allows the kernels execution commands to be sent to the device. This effectively queues them for execution, this en-queuing can be in-order or out-of-order.
+	\item{Memory objects}\newline
+	The memory objects are what is provided to the kernels.
+Kernels do accept basic types like floats, ints, bools etc.
+and self defined types like Ray, Intersection etc, but if there is a large buffer of those it's kept in a Memory buffer (e.g.
+images).
+\end{itemize}
+
+\pagebreak
+
+\begin{minipage}{0.6\textwidth}
+	\centering
+	\includegraphics[width=\textwidth]{memorymodel.png}
+\end{minipage} \hfill
+\begin{minipage}{0.4\textwidth}
+	A large part of the parallelization control is in the hands of the programmer.
+Upon launching a kernel the programmer is in control of how many work items need to be handled per compute unit, this effectively bundles them in a workgroup.
+Each workgroup gets its own compute unit, a work group can exist out of one or multiple work items.
+This degree of freedom is further reinforced by the level of control over different memory levels that can be employed to speed up development.
+\end{minipage}
+
+\vspace{0.25cm}
+
+Under this section is a roundup of the OpenCL memory model (GPU memory model).
+
+\begin{itemize}
+	\item{Global memory}\newline
+	Global memory is very similar to RAM, it is the largest memory buffer but also the slowest. Global memory can be written to / read from by all work items (or threads).
+	\item{Constant memory}\newline
+	A small chunk of global memory on the device, can be read from by all work items, but can not be written to.
+However the host application has full read/write access to this memory pool.
+Constant memory is a bit faster than
+	\item{Local memory}\newline
+	The local memory is a shared memory pool among work items in the same work group, similar to cache on the CPU.
+Local memory allows work items to share their results with work items belonging in the same work group.
+Local memory is up to 100 times faster than global memory.
+	\item{Private memory}\newline
+	The fastest type of memory, similar to registers on the cpu. Each work item (thread) has a tiny buffer of private memory to store intermediate results in.
+\end{itemize}
+
+\subsubsection{Radeon Rays}
+Radeon-Rays is a GPU intersection library(raytracing framework), was developed by AMD and it is featured in their AMD ProRenderer (raytraced renderer).
+Radeon Rays exposes a well defined C++ API for scene construction and performing asynchronous ray intersection queries.
+Radeon-Rays employs OpenCL kernels to do all these calculations and to speed up the ray intersection they order their objects in bounding volume hierarchies.
+Seeing they use OpenCL to speed up their computations by sending it of to the GPU, this library will run on almost any GPU (meaning it will also work on Nvidia/Intel products).
+
+\subsection{Rasterization}
+\subsubsection{OpenGL}
+\begin{itemize}
+	\item{OpenGL}
+	\pagebreak
+	\item{Loading libraries}\newline\par
+	OpenGL is a specification that has many different versions, implementations and features.
+The hardware manufacturers have to make sure their hardware/drivers provides the graphics programmers of those features.
+However when working with immensely different target platforms, ranging from old hardware to new and from obscure cards to bleeding edge ones the graphics programmers need some kind of way to determine what's possible on that specific piece of hardware.
+That is where loading libraries come in, loading libraries essentially provide the programmer with function pointers to OpenGL agnostic functions at runtime, meaning they provide us the memory addresses of those functions seeing they depend from hardware to hardware.\newline\par
+
+	An additional task of the loading libraries is extension loading where programmers can use this API to ask if a specific extension/feature is supported on the hardware.
+Loading libraries provide their own OpenGL implementations (e.g.
+'gl.h'), but your own implementation can be provided by making sure it is set as the last include in your build order.
+A word of caution however is that some features might be declared in your specification that the loading library has no knowledge of, but this is an edge case seeing that most people won't thread that route.\newline\par
+
+	There are numerous libraries out there that do just that, libraries like GLEW, GL3W, Glad, Glee and many more.
+All libraries that exist out there all deliver on the same promise, meaning that the choice of library really boils down to personal preference.\newline\par
+	\item{GLM}\newline\par
+	When it comes to math libraries that plug in nicely with an OpenGL renderer there's really only one good option namely GLM.
+OpenGL Mathematics (GLM) is a header only C++ mathematics library for graphics software based on the OpenGL Shading Language (GLSL).
+GLM exposes a well defined interface for everything from general calculus to quaternion rotations (e.g.
+vector operations, matrix calculations).
+\end{itemize}
+\subsubsection{Window creation}
+A window allows the rendering specification to offload its framebuffer to the viewport, it also acts as a layer between the operating system and the game when it comes to input etc.
+The creation of a window is entirely dependent on the operating system it's creating the window for.
+For example when creating a window on linux the X window system should be addressed, whereas on Windows you would need to hook on to the WinMain function and use the WndProc callback to fetch any input.
+Therefore it is much smarter to use a library that hides all this operating system agnostic stuff in an abstraction layer.\newline\par
+SDL, SFML and GLFW are all low level libraries that handle multimedia indirection.
+They all deliver the same features in some shape or form.
+The choice really boils down to maturity of the library, documentation, usefulness of its features and frequency of the updates (if any).
+\subsubsection{Model loading}
+\subsubsection{Texture loading}
+\pagebreak
+
 \section{Hybrid rendering approach}
-\begin{figure}[h]
-	\includegraphics[width=\textwidth]{hybrid.jpg}
-	\caption{Hybrid rendering empowered by Nvidia RTX}
-\end{figure}
 The different rendering techniques have been covered, now onto a hybrid solution where rasterization and raytracing are combined to achieve a new pipeline.
@@
 Ray-tracing comes into play for sending out our reflected rays, which could be anything from sending out new primary rays to secondary (shadow) rays.
-This introduction of raytracing essentially brings us world information whereas normally you'd only have surface information.
+This introduction of ray-tracing essentially brings us world information whereas normally you'd only have surface information.
 To be able to augment rasterization with raytracing we will need to postpone our lighting effects to a later stage.
-In a forward rendering pipeline the lighting is done per mesh in the fragment shader stage, effectively disabling us from hooking raytracing to that pipeline.
 
-\pagebreak
 \subsection{Raytracing in rasterization}
@@
 \subsubsection{Forward vs deferred rendering}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\textwidth]{fvsdef.png}
+	\caption{Difference between forward and deferred pipeline.}
+\end{figure}
+
+In rasterization there are three main lighting pipelines.
+The default pipeline is forward rendering this is the pipeline that's discussed in the previous chapters.
+The main difference between both pipelines however is in the way they handle the lighting effects.
+In forward rendering the lighting is applied per object in the fragment shader, this means that even the transparent objects, whose pixel fragments are not sure to end up in the final image, are also being handled.
+This pipeline is fine for default drawing, but starts to fall behind when it comes to lighting performance.\newline\par
+
+\pagebreak
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\textwidth]{fvsdefgraph.png}
+	\caption{Lighting performance comparison between forward and deferred.}
+\end{figure}
+
+The second pipeline however as the name suggests defers from doing any lighting effects until all the objects are rendered in the off-screen g-buffers.
+This effectively divides the pipeline into two major parts, the fill pass and the lighting pass.
+In the fill pass we fill up all the off-screen buffers (or G-Buffers) with our pixel in formation.
+The light pass as the name suggests is then in charge of applying lighting to those pixels.
+This ensures that we can postpone our lighting to the lightpass enabling us to do additional effects that encompass the whole scene (e.g.
+raytracing.)\newline\par
+
+The third rendering pipeline being forward+ although the most performant will not be covered with explicit detail.
+The reason that this technique is not that applicable has primarily to do with the fact that we have no way of postponing lighting effects.
+This technique also covers lighting object per object and that is precisely the opposite of what raytracing is trying to overcome.\newline\par
+Forward+ is a rendering technique that improves upon regular forward rendering.
+Forward+ first determines which lights are overlapping which ares of the screen (screen space).
+During the shading phase that list is then consulted to only account for the lights that are potentially overlapping the current pixel fragment.
+The forward+ technique consists primarily of three simple passes.\newline\par
+
+\pagebreak
+
+\subsubsection{Ray construction using g-buffers}
+
+\pagebreak
+
+\subsection{Related work}
+\pagebreak
+
+%%%%%%%%%%%%%%%%
+% NEW CHAPTER! %
+%%%%%%%%%%%%%%%%
+\chapter{Case study}
+In the following chapter an example rendering pipeline will be covered.
+This pipeline will make use of deferred shading (rasterization) for its main pipeline and raytracing for real-time shadows.
+The implementation of the shadows acts as an example effect to display that there is gain to be had when replacing that with a raytraced counterpart.
+The case study will act as a logical guide on how to implement this pipeline yourself, but some situation agnostic details will be left open for interpretation.\newline\par
+
+\section{Getting started}
+\subsection{Building using CMake}
+
+\pagebreak
+
+\section{Implementation rasterization}
+\subsection{Window creation}
+The first main part that needs to be implemented to render is a window. \newline\par
+
+\subsection{Rendering specification}
+\subsubsection{OpenGL function bindings}
+
+\subsection{Shader loading}
+One of the major things that make the graphics pipeline modular is programmable shaders.
+Shaders are mostly written in a pseudo c-language like GLSL/HLSL.
+This language is relative to the underlying rendering api that is being used.
+The rendering api is therefore in charge for the shader compilation.
+The only obstacle that the host program has to overcome is loading the source from the shader file.
+In this implementation the default cpp iostream was used, but this can be tailored to the needs of your engine/implementation.
+Once the shader source has been loaded it needs to be compiled and hooked to an OpenGL object.
+
+\subsection{Coordinate systems}
 Lorem ipsum dolor sit amet, consectetur adipiscing elit.
@@
 
-\subsubsection{Deferred shading}
+\subsection{Model loading}
+Assimp as the name implies is an asset import library that handles about every relevant model export type that exist.
+The assimp api needs to be built using CMake and can then be linked into our solution.
+Then it's just a simple case of creating an instance of their importer context and calling the readfile method.
+\subsection{Texture loading}
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Duis risus ante, auctor et pulvinar non, posuere ac lacus.
+Praesent egestas nisi id metus rhoncus ac lobortis sem hendrerit.
+Etiam et sapien eget lectus interdum posuere sit amet ac urna.
+Aliquam pellentesque imperdiet erat, eget consectetur felis malesuada quis.
+Pellentesque sollicitudin, odio sed dapibus eleifend, magna sem luctus turpis, id aliquam felis dolor eu diam.
+Etiam ullamcorper, nunc a accumsan adipiscing, turpis odio bibendum erat, id convallis magna eros nec metus.
+Sed vel ligula justo, sit amet vestibulum dolor.
+Sed vitae augue sit amet magna ullamcorper suscipit.
+Quisque dictum ipsum a sapien egestas facilisis.
+\subsection{Deferred shading}
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Duis risus ante, auctor et pulvinar non, posuere ac lacus.
+Praesent egestas nisi id metus rhoncus ac lobortis sem hendrerit.
+Etiam et sapien eget lectus interdum posuere sit amet ac urna.
+Aliquam pellentesque imperdiet erat, eget consectetur felis malesuada quis.
+Pellentesque sollicitudin, odio sed dapibus eleifend, magna sem luctus turpis, id aliquam felis dolor eu diam.
+Etiam ullamcorper, nunc a accumsan adipiscing, turpis odio bibendum erat, id convallis magna eros nec metus.
+Sed vel ligula justo, sit amet vestibulum dolor.
+Sed vitae augue sit amet magna ullamcorper suscipit.
+Quisque dictum ipsum a sapien egestas facilisis.
+\section{Implementation ray-tracing}
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Duis risus ante, auctor et pulvinar non, posuere ac lacus.
+Praesent egestas nisi id metus rhoncus ac lobortis sem hendrerit.
+Etiam et sapien eget lectus interdum posuere sit amet ac urna.
+Aliquam pellentesque imperdiet erat, eget consectetur felis malesuada quis.
+Pellentesque sollicitudin, odio sed dapibus eleifend, magna sem luctus turpis, id aliquam felis dolor eu diam.
+Etiam ullamcorper, nunc a accumsan adipiscing, turpis odio bibendum erat, id convallis magna eros nec metus.
+Sed vel ligula justo, sit amet vestibulum dolor.
+Sed vitae augue sit amet magna ullamcorper suscipit.
+Quisque dictum ipsum a sapien egestas facilisis.
+\subsection{OpenCL implementation}
 Lorem ipsum dolor sit amet, consectetur adipiscing elit.
@@
 
-\subsubsection{G-Buffers}
+\subsection{Radeon rays implementation}
 Lorem ipsum dolor sit amet, consectetur adipiscing elit.
@@
 Quisque dictum ipsum a sapien egestas facilisis.
-
-\subsection{Raytraced shadows}
-\begin{wrapfigure}{r}{0.4\textwidth}
-	\includegraphics[width=0.4\textwidth]{shadows.jpg}
-\end{wrapfigure}
-For example when looking at ray-traced shadows.
-To determine if a pixel should be shaded or not a primary ray is cast from that pixel into the scene, material information is then gathered and additional secondary rays are being cast towards the relevant lights.
-If that secondary ray/shadow ray is occluded by another object then that pixel will be shaded using only the material information, if it isn't occluded however we can add the target lights information to to the shaders equation.\newline\par
-
-\subsubsection{Soft shadows}
-Test test test
-
+\subsection{Shadow implementation}
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Duis risus ante, auctor et pulvinar non, posuere ac lacus.
+Praesent egestas nisi id metus rhoncus ac lobortis sem hendrerit.
+Etiam et sapien eget lectus interdum posuere sit amet ac urna.
+Aliquam pellentesque imperdiet erat, eget consectetur felis malesuada quis.
+Pellentesque sollicitudin, odio sed dapibus eleifend, magna sem luctus turpis, id aliquam felis dolor eu diam.
+Etiam ullamcorper, nunc a accumsan adipiscing, turpis odio bibendum erat, id convallis magna eros nec metus.
+Sed vel ligula justo, sit amet vestibulum dolor.
+Sed vitae augue sit amet magna ullamcorper suscipit.
+Quisque dictum ipsum a sapien egestas facilisis.
 \subsection{Optimizations}
@@
 Pellentesque sollicitudin, odio sed dapibus eleifend, magna sem luctus turpis, id aliquam felis dolor eu diam.
-
-%%%%%%%%%%%%%%%%
-% NEW CHAPTER! %
-%%%%%%%%%%%%%%%%
-\chapter{Case study}
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-
-\section{Implementation rasterization}
-\subsection{Window creation}
-\subsection{Rendering specification}
-\subsection{Shader loading}
-\subsection{Coordinate systems}
-\subsection{Model loading}
-\subsection{Texture loading}
-\subsection{Deferred shading}
-\section{Implementation ray-tracing}
-\subsection{Radeon rays implementation}
-\subsection{Shadow implementation}
-\subsection{Optimizations}
+Etiam ullamcorper, nunc a accumsan adipiscing, turpis odio bibendum erat, id convallis magna eros nec metus.
+Sed vel ligula justo, sit amet vestibulum dolor.
+Sed vitae augue sit amet magna ullamcorper suscipit.
+Quisque dictum ipsum a sapien egestas facilisis.
 

### 2018-11-03 Paper/Paper_BillietStijn_RadeonRays.tex
Worked on paper (primarily parts about gbuffers and layout)

@@
 The programmer has to set up the OpenCL execution pipeline and write the kernels.
-Under this section you'll find a quick rundown of aforementioned execution
-model.\newline
 
@@
 
+Under this section you'll find a quick rundown of aforementioned execution 
+model.\newline\par
 \begin{itemize}
-	\item{Platform}\newline
+	\item{\textbf{Platform}}\newline\par
 	The platform that we are running this code on, so which vendor (e.g. Nvidia).
-	\item{Device}\newline
+	\item{\textbf{Device}}\newline\par
 	What device are we going to run this on could be anything from CPU, GPU to APU.
-	\item{Context}\newline
+	\item{\textbf{Context}}\newline\par
 	This is the runtime interface between the host program (run on the CPU) and the device (CPU/GPU/etc.).
 The context is in charge of managing all the OpenCL resources so programs, kernels, command queues and buffers.
-	\item{Program}\newline
+	\item{\textbf{Program}}\newline\par
 	The entire OpenCL program, this could be one or more kernels/device functions bundled together into a single program file.
-	\item{Kernel}\newline
+	\item{\textbf{Kernel}}\newline\par
 	The starting point/building blocks of a OpenCL program.
@@
 The kernels are preceded with the keyword "\_\_kernel"
-	\item{Command queue}\newline
+	\item{\textbf{Command queue}}\newline\par
 	The command queue allows the kernels execution commands to be sent to the device. This effectively queues them for execution, this en-queuing can be in-order or out-of-order.
-	\item{Memory objects}\newline
+	\item{\textbf{Memory objects}}\newline\par
 	The memory objects are what is provided to the kernels.
@@
 \end{itemize}
-
-\pagebreak
 
@@
 \begin{itemize}
-	\item{Global memory}\newline
+	\item{\textbf{Global memory}}\newline\par
 	Global memory is very similar to RAM, it is the largest memory buffer but also the slowest. Global memory can be written to / read from by all work items (or threads).
-	\item{Constant memory}\newline
+	\item{\textbf{Constant memory}}\newline\par
 	A small chunk of global memory on the device, can be read from by all work items, but can not be written to.
@@
 Constant memory is a bit faster than
-	\item{Local memory}\newline
+	\item{\textbf{Local memory}}\newline\par
 	The local memory is a shared memory pool among work items in the same work group, similar to cache on the CPU.
@@
 Local memory is up to 100 times faster than global memory.
-	\item{Private memory}\newline
+	\item{\textbf{Private memory}}\newline\par
 	The fastest type of memory, similar to registers on the cpu. Each work item (thread) has a tiny buffer of private memory to store intermediate results in.
@@
 \begin{itemize}
-	\item{OpenGL}
+	\item{\textbf{OpenGL}}
 	\pagebreak
-	\item{Loading libraries}\newline\par
+	\item{\textbf{Loading libraries}}\newline\par
 	OpenGL is a specification that has many different versions, implementations and features.
@@
 All libraries that exist out there all deliver on the same promise, meaning that the choice of library really boils down to personal preference.\newline\par
-	\item{GLM}\newline\par
+	\item{\textbf{GLM}}\newline\par
 	When it comes to math libraries that plug in nicely with an OpenGL renderer there's really only one good option namely GLM.
@@
 The choice really boils down to maturity of the library, documentation, usefulness of its features and frequency of the updates (if any).
+\subsubsection{Texture loading}
+A texture is a 2D (could be 3D or 1D) image packed with color data that is used to feed into the graphics pipeline to texturize geometry.
+Textures can be stored in a dozen of file formats, each with their own structure to store the pixel data.
+Textures are stored in a binary format and can therefore be parsed by a binary reader.
+The problem however with doing textureloading is that every format is different so the programmer would be in charge of writing a dozen different loaders to support the most basic of texture types.
+Seeing that this is rather cumbersome most people would opt for a library that does this for them.
+Again there are quite a lot of libraries out there that do this.
+\newline\par
+
+SOIL, GLi, DevIL, FreeImage and stb\_image are all texture loading libraries that promise on loading the most prevalent of types.
+\pagebreak
+
 \subsubsection{Model loading}
-\subsubsection{Texture loading}
+\begin{itemize}
+	\item{\textbf{3D file format}}\newline\par
+	A 3D file format is a specification to store information about 3D models in the form of plain text or as binary data.
+The main purpose of such format is to encode 3D model's geometry, material information and animation data.
+Seeing there are loads of formats out there, creating your own model loading library has become something of a novelty.
+Fortunately there are loads of libraries out there that load about any format you throw at it.
+	\item{\textbf{AssImp}}\newline
+
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.9\textwidth]{assimpstructure.png}
+	\end{figure}
+
+	Asset import or AssImp for short is a model loading library designed to load as many formats as possible and to unify them into a searchable structure to be used by the host application.
+It structures the entire scene into a scene object, which consecutively holds a rootnode and a collection of meshes and materials.
+That rootnote then has leads to other nodes that all hold the same information namely a list of indices to the corresponding meshes and a next connection.
+This leads to a clear distinction as to what are seperate objects in the global model file, which can then later be used by the game-engine to ensure perfect relationships between aforementioned meshes.
+\end{itemize}
+
 \pagebreak
@@
 
-The third rendering pipeline being forward+ although the most performant will not be covered with explicit detail.
-The reason that this technique is not that applicable has primarily to do with the fact that we have no way of postponing lighting effects.
-This technique also covers lighting object per object and that is precisely the opposite of what raytracing is trying to overcome.\newline\par
+The third rendering pipeline, although the most performant will not be covered with explicit detail.
+The reason that this technique is not that applicable has primarily to do with the fact that we have no way of doing per world space lighting effects.
 Forward+ is a rendering technique that improves upon regular forward rendering.
-Forward+ first determines which lights are overlapping which ares of the screen (screen space).
+Forward+ first determines which lights are overlapping which areas of the screen (screen space).
 During the shading phase that list is then consulted to only account for the lights that are potentially overlapping the current pixel fragment.
-The forward+ technique consists primarily of three simple passes.\newline\par
+The forward+ technique consists primarily of three simple passes.
+In the light culling pass, each light in the scene is sorted into screen space tiles.
+In the opaque pass, the light list generated from the light culling pass is used to compute the lighting for opaque geometry.
+The transparent pass is similar to the opaque pass in the sense that the light list is again consulted when shading the transparent pixel fragments.
 
@@
 
-\subsubsection{Ray construction using g-buffers}
+\subsubsection{Raytracing using g-buffers}
 
-\pagebreak
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\textwidth]{gbuffer.png}
+	\caption{Illustration of g-buffers}
+\end{figure}
 
-\subsection{Related work}
+As touched earlier deferred rendering fills the off screen render buffers (g-buffers) with scene information in its first pass.
+This effectively enables us to do rendering after the fact by then using that information to construct everything we need to apply our lighting.\newline\par
+
+The information that is stored in the g-buffers is entirely up to the programmer (e.g.
+position data, normal data, color data etc.).
+However to be able to use this in conjunction with ray construction the programmer needs atleast the position data and the normal data.
+Some people would however argue that the position data isn't necessary and that the position can be calculated by the z-buffer and that would be an entirely correct answer, but for simplicity sake people opt for the position data instead.\newline\par
+
+With the geometry data now in the g-buffers, that data can now be used to construct the rays.
+The ray construction can be sped up by doing it on the GPU, seeing that this operation is highly parallel in nature.
+Essentially what it boils down to is to make sure that meshes have been synchronized with the raytracer, meaning that the raytracer has all the right meshes in it's scene to be able to fetch all occlusion/intersections.
+Then quite simply sample all the position entries and normal entries (for the entire screen) and send out a ray for every pixel, this will in turn correlate to the raytracer api seeing those position entries should correspond with the meshes in the scene.\newline\par
+
+Once the rays have been constructed the raytracer can now simply query the occlusion/intersections with the scene geometry. That data can then be used to shade the object respectively.\newline\par
+
 \pagebreak
@@
 \subsection{Coordinate systems}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Duis risus ante, auctor et pulvinar non, posuere ac lacus.
-Praesent egestas nisi id metus rhoncus ac lobortis sem hendrerit.
-Etiam et sapien eget lectus interdum posuere sit amet ac urna.
-Aliquam pellentesque imperdiet erat, eget consectetur felis malesuada quis.
-Pellentesque sollicitudin, odio sed dapibus eleifend, magna sem luctus turpis, id aliquam felis dolor eu diam.
-Etiam ullamcorper, nunc a accumsan adipiscing, turpis odio bibendum erat, id convallis magna eros nec metus.
-Sed vel ligula justo, sit amet vestibulum dolor.
-Sed vitae augue sit amet magna ullamcorper suscipit.
-Quisque dictum ipsum a sapien egestas facilisis.
 
 \subsection{Model loading}
-Assimp as the name implies is an asset import library that handles about every relevant model export type that exist.
-The assimp api needs to be built using CMake and can then be linked into our solution.
-Then it's just a simple case of creating an instance of their importer context and calling the readfile method.
+
 \subsection{Texture loading}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Duis risus ante, auctor et pulvinar non, posuere ac lacus.
-Praesent egestas nisi id metus rhoncus ac lobortis sem hendrerit.
-Etiam et sapien eget lectus interdum posuere sit amet ac urna.
-Aliquam pellentesque imperdiet erat, eget consectetur felis malesuada quis.
-Pellentesque sollicitudin, odio sed dapibus eleifend, magna sem luctus turpis, id aliquam felis dolor eu diam.
-Etiam ullamcorper, nunc a accumsan adipiscing, turpis odio bibendum erat, id convallis magna eros nec metus.
-Sed vel ligula justo, sit amet vestibulum dolor.
-Sed vitae augue sit amet magna ullamcorper suscipit.
-Quisque dictum ipsum a sapien egestas facilisis.
+
 \subsection{Deferred shading}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Duis risus ante, auctor et pulvinar non, posuere ac lacus.
-Praesent egestas nisi id metus rhoncus ac lobortis sem hendrerit.
-Etiam et sapien eget lectus interdum posuere sit amet ac urna.
-Aliquam pellentesque imperdiet erat, eget consectetur felis malesuada quis.
-Pellentesque sollicitudin, odio sed dapibus eleifend, magna sem luctus turpis, id aliquam felis dolor eu diam.
-Etiam ullamcorper, nunc a accumsan adipiscing, turpis odio bibendum erat, id convallis magna eros nec metus.
-Sed vel ligula justo, sit amet vestibulum dolor.
-Sed vitae augue sit amet magna ullamcorper suscipit.
-Quisque dictum ipsum a sapien egestas facilisis.
+
 \section{Implementation ray-tracing}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Duis risus ante, auctor et pulvinar non, posuere ac lacus.
-Praesent egestas nisi id metus rhoncus ac lobortis sem hendrerit.
-Etiam et sapien eget lectus interdum posuere sit amet ac urna.
-Aliquam pellentesque imperdiet erat, eget consectetur felis malesuada quis.
-Pellentesque sollicitudin, odio sed dapibus eleifend, magna sem luctus turpis, id aliquam felis dolor eu diam.
-Etiam ullamcorper, nunc a accumsan adipiscing, turpis odio bibendum erat, id convallis magna eros nec metus.
-Sed vel ligula justo, sit amet vestibulum dolor.
-Sed vitae augue sit amet magna ullamcorper suscipit.
-Quisque dictum ipsum a sapien egestas facilisis.
+
 \subsection{OpenCL implementation}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Duis risus ante, auctor et pulvinar non, posuere ac lacus.
-Praesent egestas nisi id metus rhoncus ac lobortis sem hendrerit.
-Etiam et sapien eget lectus interdum posuere sit amet ac urna.
-Aliquam pellentesque imperdiet erat, eget consectetur felis malesuada quis.
-Pellentesque sollicitudin, odio sed dapibus eleifend, magna sem luctus turpis, id aliquam felis dolor eu diam.
-Etiam ullamcorper, nunc a accumsan adipiscing, turpis odio bibendum erat, id convallis magna eros nec metus.
-Sed vel ligula justo, sit amet vestibulum dolor.
-Sed vitae augue sit amet magna ullamcorper suscipit.
-Quisque dictum ipsum a sapien egestas facilisis.
 
 \subsection{Radeon rays implementation}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Duis risus ante, auctor et pulvinar non, posuere ac lacus.
-Praesent egestas nisi id metus rhoncus ac lobortis sem hendrerit.
-Etiam et sapien eget lectus interdum posuere sit amet ac urna.
-Aliquam pellentesque imperdiet erat, eget consectetur felis malesuada quis.
-Pellentesque sollicitudin, odio sed dapibus eleifend, magna sem luctus turpis, id aliquam felis dolor eu diam.
-Etiam ullamcorper, nunc a accumsan adipiscing, turpis odio bibendum erat, id convallis magna eros nec metus.
-Sed vel ligula justo, sit amet vestibulum dolor.
-Sed vitae augue sit amet magna ullamcorper suscipit.
-Quisque dictum ipsum a sapien egestas facilisis.
+
 \subsection{Shadow implementation}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Duis risus ante, auctor et pulvinar non, posuere ac lacus.
-Praesent egestas nisi id metus rhoncus ac lobortis sem hendrerit.
-Etiam et sapien eget lectus interdum posuere sit amet ac urna.
-Aliquam pellentesque imperdiet erat, eget consectetur felis malesuada quis.
-Pellentesque sollicitudin, odio sed dapibus eleifend, magna sem luctus turpis, id aliquam felis dolor eu diam.
-Etiam ullamcorper, nunc a accumsan adipiscing, turpis odio bibendum erat, id convallis magna eros nec metus.
-Sed vel ligula justo, sit amet vestibulum dolor.
-Sed vitae augue sit amet magna ullamcorper suscipit.
-Quisque dictum ipsum a sapien egestas facilisis.
+
 \subsection{Optimizations}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Duis risus ante, auctor et pulvinar non, posuere ac lacus.
-Praesent egestas nisi id metus rhoncus ac lobortis sem hendrerit.
-Etiam et sapien eget lectus interdum posuere sit amet ac urna.
-Aliquam pellentesque imperdiet erat, eget consectetur felis malesuada quis.
-Pellentesque sollicitudin, odio sed dapibus eleifend, magna sem luctus turpis, id aliquam felis dolor eu diam.
-Etiam ullamcorper, nunc a accumsan adipiscing, turpis odio bibendum erat, id convallis magna eros nec metus.
-Sed vel ligula justo, sit amet vestibulum dolor.
-Sed vitae augue sit amet magna ullamcorper suscipit.
-Quisque dictum ipsum a sapien egestas facilisis.
 
-\chapter{Conlusion}
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+\chapter{Conclusion}
 

### 2018-11-03 Paper/Paper_BillietStijn_RadeonRays.tex
Fixed small typo, on to case study

@@
 As touched earlier deferred rendering fills the off screen render buffers (g-buffers) with scene information in its first pass.
-This effectively enables us to do rendering after the fact by then using that information to construct everything we need to apply our lighting.\newline\par
+This effectively enables us to do lighting after the fact by then using that information to construct everything we need to apply our lighting.\newline\par
 

### 2018-11-03 Paper/Paper_BillietStijn_RadeonRays.tex
Fixed layout issue :D

@@
 \end{figure}
-
+Under this section you'll find a quick rundown of aforementioned execution 
+model.
 \pagebreak
-
-Under this section you'll find a quick rundown of aforementioned execution 
-model.\newline\par
 \begin{itemize}
@@
 
-\begin{minipage}{0.6\textwidth}
-	\centering
-	\includegraphics[width=\textwidth]{memorymodel.png}
-\end{minipage} \hfill
-\begin{minipage}{0.4\textwidth}
 A large part of the parallelization control is in the hands of the programmer.
@@
 Each workgroup gets its own compute unit, a work group can exist out of one or multiple work items.
-This degree of freedom is further reinforced by the level of control over different memory levels that can be employed to speed up development.
-\end{minipage}
+This degree of freedom is further reinforced by the level of control over different memory levels that can be employed to speed up development.\newline\par
 
-\vspace{0.25cm}
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.8\textwidth]{memorymodel.png}
+	\caption{Illustration of the memory execution model.}
+\end{figure}
 
-Under this section is a roundup of the OpenCL memory model (GPU memory model).
+Under this section is a roundup of the OpenCL memory model (GPU memory model).\newline\par
+\pagebreak
 
@@
 Radeon-Rays is a GPU intersection library(raytracing framework), was developed by AMD and it is featured in their AMD ProRenderer (raytraced renderer).
-Radeon Rays exposes a well defined C++ API for scene construction and performing asynchronous ray intersection queries.
+Radeon Rays exposes a well defined C++ API for scene construction and performing asynchronous ray intersection queries.\newline\par
 Radeon-Rays employs OpenCL kernels to do all these calculations and to speed up the ray intersection they order their objects in bounding volume hierarchies.

### 2018-11-03 Paper/Paper_BillietStijn_RadeonRays.tex
Worked on openCL interop part

@@
 
+\pagebreak
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\textwidth]{sharegroups.png}
+	\caption{OpenCL/OpenGL interop context}
+\end{figure}
+
+A great thing about this control of memory is that it empowers the developer to make use of interoperation contexts.
+This can be used to achieve zero-copy architectures where memory objects can be shared by multiple compatible api's.
+A prime example of this is an OpenCL/OpenGL interop context.
+Such a context allows the graphics programmers to send screen buffers, textures and openGL objects to the kernels for further processing without the need to copy them, further enabling them to use a kernel in the pipeline everywhere they want.\newline\par
+
+While this is great, there are still some limitations that have to be set in place to enable this kind of behaviour.
+For one the target device needs to support this sharegroup behaviour and for seconds shared memory can cause issues seeing that both contexts don't know about each others existance.
+All of this can be alliviated by setting blocking calls (e.g.
+GPU fences) in place, with these acting as a sort of failsafe that makes sure the object is done reading from/ writing to by the target context.
+This can account for slowdowns seeing that the CPU is needed to help the 2 parties syncronize, luckily for the programmer there is hope that said GPU supports the ARB\_cl\_event extension.3 The event extension allows the two api's to communicate through event management (e.g.
+wait for event to begin) ensuring that they don't need to wait a few ticks of CPU-time to synchronize.
+
+\pagebreak
+
 \subsubsection{Radeon Rays}

### 2018-11-03 Paper/Paper_BillietStijn_RadeonRays.tex
Fixed a load of typos (many more to come i'm afraid).

@@
 The results of rasterization have gotten very good over the years, very convincing at relatively low compute performance.
-However there are certain limitations to this algorithm, seeing it's main goal is to get graphics in screen space.\newline\par
+However there are certain limitations to this algorithm, seeing its main goal is to get graphics in screen space.\newline\par
 
@@
 \subsubsection{Rasterization stage}
-The fourth stage is rasterization, this is where the algorithm gets it's name.
+The fourth stage is rasterization, this is where the algorithm gets its name.
 The rasterizer is in charge of a not only rasterization, but also homogeneous divide, clipping, culling and viewport transform.
@@
 DirectX/OpenGL) rasterizer.
-In it's essence the calculation will translate to if the dotproduct is negative then the face is facing away, if it's positive then it's facing toward us.
+In its essence the calculation will translate to if the dotproduct is negative then the face is facing away, if it's positive then it's facing toward us.
 \end{minipage}
@@
 
-A great thing about this control of memory is that it empowers the developer to make use of interoperation contexts.
+A great thing about this control over memory is that it empowers the developer to make use of interoperation contexts.
 This can be used to achieve zero-copy architectures where memory objects can be shared by multiple compatible api's.
@@
 GPU fences) in place, with these acting as a sort of failsafe that makes sure the object is done reading from/ writing to by the target context.
-This can account for slowdowns seeing that the CPU is needed to help the 2 parties syncronize, luckily for the programmer there is hope that said GPU supports the ARB\_cl\_event extension.3 The event extension allows the two api's to communicate through event management (e.g.
+This can account for slowdowns seeing that the CPU is needed to help the 2 parties synchronize, luckily for the programmer there is hope that said GPU supports the ARB\_cl\_event extension.3 The event extension allows the two api's to communicate through event management (e.g.
 wait for event to begin) ensuring that they don't need to wait a few ticks of CPU-time to synchronize.
@@
 OpenGL Mathematics (GLM) is a header only C++ mathematics library for graphics software based on the OpenGL Shading Language (GLSL).
-GLM exposes a well defined interface for everything from general calculus to quaternion rotations (e.g.
-vector operations, matrix calculations).
+\newline\par
+	GLM exposes a well defined interface for everything from calculus to quaternion rotations (e.g. vector operations, matrix calculations).
 \end{itemize}
@@
 The third rendering pipeline, although the most performant will not be covered with explicit detail.
-The reason that this technique is not that applicable has primarily to do with the fact that we have no way of doing per world space lighting effects.
+The reason that this technique is not that applicable has primarily to do with the fact that we have no way of doing world space lighting effects.
 Forward+ is a rendering technique that improves upon regular forward rendering.
@@
 The ray construction can be sped up by doing it on the GPU, seeing that this operation is highly parallel in nature.
-Essentially what it boils down to is to make sure that meshes have been synchronized with the raytracer, meaning that the raytracer has all the right meshes in it's scene to be able to fetch all occlusion/intersections.
+Essentially what it boils down to is to make sure that meshes have been synchronized with the raytracer, meaning that the raytracer has all the right meshes in its scene to be able to fetch all occlusion/intersections.
 Then quite simply sample all the position entries and normal entries (for the entire screen) and send out a ray for every pixel, this will in turn correlate to the raytracer api seeing those position entries should correspond with the meshes in the scene.\newline\par

### 2018-11-03 Paper/Paper_BillietStijn_RadeonRays.tex
Worked on first parts of case study

@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{memorymodel.png}
+	\includegraphics[width=0.7\textwidth]{memorymodel.png}
 	\caption{Illustration of the memory execution model.}
@@
 \subsection{Window creation}
-The first main part that needs to be implemented to render is a window. \newline\par
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\textwidth]{window.jpg}
+	\caption{Snippet for window creation}
+\end{figure}
+
+The first main part that needs to be implemented to render is a window.
+As already outlined in the research part there are a lot of libraries out there.
+For this implementation SDL was chosen, simply for the reason that creating an SDL windowing context with OpenGL is a breeze.
+SDL expects the programmer to supply what version of OpenGL will be coupled to this window, what rendering mode will be used (e.g.
+immediate vs core) and the settings of the window.
+In our case we'll use OpenGL 3.3 seeing it was the first modern implementation of SDL and consecutively the easiest to comprehend, we'll also be using core mode seeing we need vertex buffer abilities.\newline\par
 
@@
 \subsubsection{OpenGL function bindings}
+Once the window and accompanying context is in place the programmer can move on to the OpenGL loading.
+For this implementation Glad was used.
+To get glad to work in this solution the respective package has to be created from their website \href{https://glad.dav1d.de/}.
+For our purpose I selected the GL 3.3 package in core functionality mode.
+After the package selection has been done you should receive the source files.
+These can then be compiled into a solution by making use of CMake, to then consecutively build those into lib files.
+The final step was to link both the includes and libs into my solution.
+A simple call to gladLoadGLLoader(SDL\_GL\_GetProcAddress) later and it's done.\newline\par
 
+\pagebreak
 \subsection{Shader loading}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\textwidth]{shader.jpg}
+\end{figure}
+
 One of the major things that make the graphics pipeline modular is programmable shaders.
@@
 Once the shader source has been loaded it needs to be compiled and hooked to an OpenGL object.
+In this implementation a clear distinction between a shader and a shader program was made, where a shader is seen as the smallest building block and a shader program as a collection of those blocks.\newline\par
 

### 2018-11-03 Paper/Paper_BillietStijn_RadeonRays.tex
Worked on paper (Case study)

@@
 
-\subsection{Coordinate systems}
-
 \subsection{Model loading}
@@
 \subsection{Texture loading}
+
+\subsection{Coordinate systems}
+Although you can render without a proper coordinate system in place, the implementation needed world space coordinates rather than screen space. \newline\par
+So for this implementation a transform component was set in place.
+This transform component stands in relationship with the GameObject.
+This relationship allows us to add childs to the GameObject while maintaining the correct local transforms in the component.
+If the position of a said object was changed we go ahead and rebuild our worldmatrix while keeping in check if the object had a parent.\newline\par
+The next part is the camera component, the camera component is in charge of holding the viewprojectionmatrix.
+Meaning that it is also in charge of holding the camera position and its settings.\newline\par
+The final part to the puzzle is the mesh draw component.
+To create a meshdrawcomponent the programmer has to provide it with a mesh filter which is loaded by the model loading utilities mentioned prior.
+Then the final assembly is to build the worldviewprojectionmatrix (by multiplying the world, view and projection matrices with each other) and send it to the shader.
 
@@
 \subsection{Shadow implementation}
-
-\subsection{Optimizations}
 

### 2018-11-03 Paper/Paper_BillietStijn_RadeonRays.tex
Worked on case study paper (Rasterization part)

@@
 	\includegraphics[width=\textwidth]{window.jpg}
-	\caption{Snippet for window creation}
 \end{figure}
@@
 immediate vs core) and the settings of the window.
-In our case we'll use OpenGL 3.3 seeing it was the first modern implementation of SDL and consecutively the easiest to comprehend, we'll also be using core mode seeing we need vertex buffer abilities.\newline\par
+In our case we'll use OpenGL 3.3 seeing it was the first modern implementation of SDL and consecutively the easiest to comprehend, we'll also be using core mode seeing we need vertex buffer abilities.
 
-\subsection{Rendering specification}
-\subsubsection{OpenGL function bindings}
+\subsection{OpenGL rendering context}
+Once the window is in place the programmer is responsible for querying the accompanying OpenGL context.
+
+\subsection{OpenGL function bindings}
 Once the window and accompanying context is in place the programmer can move on to the OpenGL loading.
@@
 
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=\textwidth]{shader.jpg}
-\end{figure}
-
 One of the major things that make the graphics pipeline modular is programmable shaders.
@@
 Once the shader source has been loaded it needs to be compiled and hooked to an OpenGL object.
-In this implementation a clear distinction between a shader and a shader program was made, where a shader is seen as the smallest building block and a shader program as a collection of those blocks.\newline\par
+In this implementation a clear distinction between a shader and a shader program was made, where a shader is seen as the smallest building block and a shader program as a collection of those blocks.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\textwidth]{shader.jpg}
+	\caption{Snippet from shader loader}
+\end{figure}
 
 \subsection{Model loading}
+As mentioned prior for this implementation AssImp was used.
+AssImp provides a very simple interface only requiring us programmers to call the importers read file and to store it into a aiScene object.
+This object is essentially our top level node that allows us to use this tree as a searchable structure.
+To keep things easy our implementation does only require us to make 3 functions being: ProcessNode, ProcessMesh and LoadMaterialTextures.
+Under this section you'll find a quick rundown of what each function does and why.
+
+\begin{itemize}
+	\item{ProcessNode}\newline\par
+	ProcessNode as the name suggest goes over the current node and recursively calls the ProcessMesh on its meshes.
+Once all the meshes within this node have been handled it goes and calls ProcessNode on its children.
+	\item{ProcessMesh}\newline\par
+	ProcessMesh has three major tasks the first one being fetching all the vertices of the mesh things like: storing position, normal data and texture coordinates.
+Secondly it makes sure all the indices are being loaded in before finally going over to the materials where it calls the LoadMaterialTextures function on each of its materials.
+	\item{LoadMaterialTextures}\newline\par
+	Finally LoadMaterialTextures then takes in the material and the type of texture that we are trying to fetch, it then checks the 3D object format for that type of format and stores it respectively.
+As an added optimization it keeps track of double textures, so it won't load the same texture twice.
+\end{itemize}
+
+\pagebreak
 
 \subsection{Texture loading}
+For texture loading stb\_image was used to accelerate this task.
+stb\_image is a very simple library and is include file only, so there is not need to create a library file.
+stb\_image provides the programmer with a very simple api for loading textures and it is up to the programmer to link those to openGL objects.
 
@@
 
+\pagebreak
+
 \subsection{Deferred shading}
+
+\pagebreak
 

### 2018-11-03 Paper/Paper_BillietStijn_RadeonRays.tex
Worked on case study part of my paper

@@
 \subsubsection{OpenGL}
-\begin{itemize}
-	\item{\textbf{OpenGL}}
 \pagebreak
-	\item{\textbf{Loading libraries}}\newline\par
+\subsubsection{Loading libraries}
 	OpenGL is a specification that has many different versions, implementations and features.
@@
 	There are numerous libraries out there that do just that, libraries like GLEW, GL3W, Glad, Glee and many more.
-All libraries that exist out there all deliver on the same promise, meaning that the choice of library really boils down to personal preference.\newline\par
-	\item{\textbf{GLM}}\newline\par
+All libraries that exist out there all deliver on the same promise, meaning that the choice of library really boils down to personal preference.
+\subsubsection{GLM}
 	When it comes to math libraries that plug in nicely with an OpenGL renderer there's really only one good option namely GLM.
@@
 	GLM exposes a well defined interface for everything from calculus to quaternion rotations (e.g. vector operations, matrix calculations).
-\end{itemize}
+\pagebreak
 \subsubsection{Window creation}
@@
 \subsubsection{Model loading}
-\begin{itemize}
-	\item{\textbf{3D file format}}\newline\par
 	A 3D file format is a specification to store information about 3D models in the form of plain text or as binary data.
@@
 Fortunately there are loads of libraries out there that load about any format you throw at it.
-	\item{\textbf{AssImp}}\newline
 
@@
 This leads to a clear distinction as to what are seperate objects in the global model file, which can then later be used by the game-engine to ensure perfect relationships between aforementioned meshes.
-\end{itemize}
 

### 2018-11-04 Paper/Paper_BillietStijn_RadeonRays.tex
Worked on paper (only one major piece left)

@@
 
-\section{Getting started}
-\subsection{Building using CMake}
-
 \pagebreak
@@
 In our case we'll use OpenGL 3.3 seeing it was the first modern implementation of SDL and consecutively the easiest to comprehend, we'll also be using core mode seeing we need vertex buffer abilities.
-
-\subsection{OpenGL rendering context}
 Once the window is in place the programmer is responsible for querying the accompanying OpenGL context.
@@
 For this implementation Glad was used.
-To get glad to work in this solution the respective package has to be created from their website \href{https://glad.dav1d.de/}.
-For our purpose I selected the GL 3.3 package in core functionality mode.
+To get glad to work in this solution the respective package has to be created from their website \url{https://glad.dav1d.de/} For our purpose I selected the GL 3.3 package in core functionality mode.
 After the package selection has been done you should receive the source files.
@@
 \pagebreak
+
 \subsection{Shader loading}
-
 One of the major things that make the graphics pipeline modular is programmable shaders.
@@
 
+\subsection{Hello Triangle}
+With OpenGL and SDL established it is time to test the renderer by doing to famous hello triangle test.
+The first step for the programmer is to supply the vertices of the triangle, seeing there are still no coordinate systems in place it is ok to define those vertices in normalized device coordinates.
+
 \begin{figure}[h]
 	\centering
-	\includegraphics[width=\textwidth]{shader.jpg}
-	\caption{Snippet from shader loader}
+	\includegraphics[width=0.6\textwidth]{ndc.png}
+	\caption{Normalized device coordinates for our triangle}
 \end{figure}
 
-\subsection{Model loading}
-As mentioned prior for this implementation AssImp was used.
-AssImp provides a very simple interface only requiring us programmers to call the importers read file and to store it into a aiScene object.
-This object is essentially our top level node that allows us to use this tree as a searchable structure.
-To keep things easy our implementation does only require us to make 3 functions being: ProcessNode, ProcessMesh and LoadMaterialTextures.
-Under this section you'll find a quick rundown of what each function does and why.
-
-\begin{itemize}
-	\item{ProcessNode}\newline\par
-	ProcessNode as the name suggest goes over the current node and recursively calls the ProcessMesh on its meshes.
-Once all the meshes within this node have been handled it goes and calls ProcessNode on its children.
-	\item{ProcessMesh}\newline\par
-	ProcessMesh has three major tasks the first one being fetching all the vertices of the mesh things like: storing position, normal data and texture coordinates.
-Secondly it makes sure all the indices are being loaded in before finally going over to the materials where it calls the LoadMaterialTextures function on each of its materials.
-	\item{LoadMaterialTextures}\newline\par
-	Finally LoadMaterialTextures then takes in the material and the type of texture that we are trying to fetch, it then checks the 3D object format for that type of format and stores it respectively.
-As an added optimization it keeps track of double textures, so it won't load the same texture twice.
-\end{itemize}
+Once the vertex data is defined the programmer needs to send it of to the first stage of the rendering pipeline (vertex shader).
+This means that the programmer should first get those vertices on the GPU, in modern OpenGL this is done by requesting a piece of memory called a vertex buffer object.
+After binding this buffer and providing it from its buffer data the object should be set in place.\newline\par
 
@@
 
-\subsection{Texture loading}
-For texture loading stb\_image was used to accelerate this task.
-stb\_image is a very simple library and is include file only, so there is not need to create a library file.
-stb\_image provides the programmer with a very simple api for loading textures and it is up to the programmer to link those to openGL objects.
+Modern OpenGL requires the graphics programmer to atleast supply his own vertex and fragment shader, this is in contrary to DirectX where a default one is used if none supplied.
+Our vertex shader will consist of the vertex layout and the passing of our vertices position to the OpenGL system value gl\_Position.
+Under vertex layout the programmer provides how his vertices store the data needed in the shader, this also needs to be done in the host program through the use of glVertexAttributePointer.
+The aforementioned gl\_Position is there to provide a means of communication to specify where the vertex should appear on screen.
 
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\textwidth]{vertexattributepointer.png}
+\end{figure}
+
+Then the final piece of the puzzle is to supply our pipeline of a fragment shader.
+The fragment shader in itself is quite simple only requiring us to specify the final fragments color, this can of course be any arbitrary value.
+After shader compilation, binding our shader, binding our vertex buffer and a final call to drawtriangles later a triangle appeared on screen.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\textwidth]{hellotriangle.png}
+\end{figure}
+
+\pagebreak
 \subsection{Coordinate systems}
@@
 
+\subsection{Texture loading}
+For texture loading stb\_image was used to accelerate this task.
+stb\_image is a very simple library and is include file only, so there is not need to create a library file.
+stb\_image provides the programmer with a very simple api for loading textures and it is up to the programmer to link those to openGL objects.
+
+\pagebreak
+
+\subsection{Model loading}
+As mentioned prior for this implementation AssImp was used.
+AssImp provides a very simple interface only requiring us programmers to call the importers read file and to store it into a aiScene object.
+This object is essentially our top level node that allows us to use this tree as a searchable structure.
+To keep things easy our implementation does only require us to make 3 functions being: ProcessNode, ProcessMesh and LoadMaterialTextures.
+Under this section you'll find a quick rundown of what each function does and why.
+
+\begin{itemize}
+	\item{ProcessNode} as the name suggest goes over the current node and recursively calls the ProcessMesh on its meshes.
+Once all the meshes within this node have been handled it goes and calls ProcessNode on its children.
+	\item{ProcessMesh} has three major tasks the first one being fetching all the vertices of the mesh things like: storing position, normal data and texture coordinates.
+Secondly it makes sure all the indices are being loaded in before finally going over to the materials where it calls the LoadMaterialTextures function on each of its materials.
+	\item{LoadMaterialTextures} then takes in the material and the type of texture that we are trying to fetch, it then checks the 3D object format for that type of format and stores it respectively.
+As an added optimization it keeps track of double textures, so it won't load the same texture twice.
+\end{itemize}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.9\textwidth]{modelandtexture.png}
+\end{figure}
+
 \pagebreak
@@
 \subsection{Deferred shading}
+With a standard rendering pipeline in place it was time to go ahead and turn that pipeline into a deferred one.
+The steps beforehand did also not cover lighting effects for the simple reason that they would need to be re-written once going deferred.\newline\par
+The major implementation difference now is that the programmer needs to define his/her off screen render buffers (e.g.
+position buffer, normal buffer, etc.).
+First the programmer has to define what buffers are stored in the gbuffer object (frame buffer object).
+With the gbuffer object defined it's time to start moving on to the fill shader.\newline\par
+The fill shader is tasked with filling the gbuffers with all the geometry information respectively, so if the implementation uses four buffers the fill shader will be tasked with filling those four buffers.
+The fill shader is a shader program consisting of two shaders first a simple vertex shader that passes the information to the next stages and then a fragment shader who does all the heavy lifting and fills all the buffers.
+Then all the geometry gets rendered using this shader.
+After the bulk of the rendering is done we unbind the gbuffers (that are now filled with geometry information) and we move on to the lightpass.\newline\par
+
+The light pass shader is in charge of lighting and rendering the final full screen quad.
+It speaks for itself that before going into this lightpass we need to make sure the g-buffer is no longer bound as the active rendertarget and that the g-buffer textures can now be sent off to the lightpass to do its lighting calculations.
+Here again the implementation makes use of two shaders a very simple vertex shader that effectively passes texture coordinates and vertex position to the fragment stage.
+The fragment shader in this case will now resemble a traditional shader apart from the fact that it needs to sample all of its information from the geometry render buffers (g-buffer).\newline\par
+
+The final piece of code is in charge of rendering a full screen quad with the combined g-buffer information (through the use of the light pass) as its texture.
 

### 2018-11-05 Paper/Paper_BillietStijn_RadeonRays.tex
Worked on paper and am moving to push soft shadows

@@
 \section{Implementation ray-tracing}
+\subsection{OpenCL implementation}
+Setting up OpenCL can be a tedious process seeing there are multiple levels that need to be set up.
+For this implementation a small toolkit CLW was used.
+All this toolkit does is hide some of the abstractions behind very simple api calls.
+Nontheless it does take some freedom away from the programmer, but it does allow the programmer to jerryrig its own implementation to theirs.\newline\par
 
-\subsection{OpenCL implementation}
+As outlined in the research part there is a lot of performance to be had by just simply creating the right context through OpenCL.
+Seeing that our pipeline has three major parts that need communicating with each other (e.g.
+OpenCL/OpenGL/RadeonRays).
+To benefit from almost free communication (e.g.
+zero copy) the programmer first has to make an interop context.\newline\par
+
+The first steps for setting up OpenCL are getting the devices and platforms.
+Once the devices have been fetched and after verifying that a valid device was found the next logical step would be to create the context.
+An additional step to the context creation call is to specify context properties, now here is where it gets exponentially more interesting seeing that those properties allow the programmer to set up his interop context.
 
 \subsection{Radeon rays implementation}
+With OpenCL now set in place it is time to move on to Radeon Rays.
+To build RadeonRays the programmer can simply download it from \url{https://github.com/GPUOpen-LibrariesAndSDKs/RadeonRays_SDK} and build the binaries using CMake.
+Like addressed in the research portion of this paper Radeon Rays uses a OpenCL context to run its intersection kernels, so when initializing Radeon Rays it automatically creates a new OpenCL context.
+Luckily though there is an overloaded constructor taking an already existing OpenCL context, further enabling the programmer to share resources between OpenCL and Radeon Rays.
+
+\pagebreak
 
 \subsection{Shadow implementation}
+The shadow implementation will heavily benefit from GPU acceleration seeing that the bulk of these operations are very parallel in nature.
+\subsubsection{Ray creation}
+First the GPU will be used to create a shadow ray buffer, so it will be tasked with creating the 2073600 rays (ray for every pixel in 1920x1080p resolution).
+As you can probably tell this is an enormous amount of data, but each ray has no knowledge of the other one existing making it perfect for parallel execution.\newline\par
+
+Here the power of OpenCL and deferred rendering are combined to run a kernel for every entry in the position buffer and normal buffer.
+Effectively creating a ray that travels from the objects world position towards the lights (in this case directional light).
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.8\textwidth]{GenShadowRays.jpg}
+	\caption{Code snippet from generate shadow rays kernel}
+\end{figure}
+
+\pagebreak
+
+\subsubsection{Light mask}
+Before moving on to the light mask the programmer needs to send the ray buffer of to RadeonRays where its kernels will do the ray intersection.
+Once the intersection query has taken place the library provides us from an intersection buffer.
+This buffer is stored in GPU memory and can thus be shared with our own OpenCL kernels, alleviating the need to copy from CPU to GPU.\newline\par
+
+Again for every pixel in the light buffer, which is now white, we run a light mask kernel.
+The light mask kernel is in charge of fetching the intersection results and checking if they reached the light or not.
+For hard shadows this is easy, when the pixel gets occluded the light mask gets filled with a zero, if the ray reached the light the value stays one.
+For lighting calculations this means that pixels that have a one will get lit and pixel that have a zero will not get lit.
+This isn't totally accurate seeing we'd need to rely on global illumination to account for light bleeding.\newline\par
+
+It gets a little more interesting when looking at soft shadows however.
+Traditional ray tracing relies on multiple rays per light source to create soft shadow gradients, like covered in the research part this wouldn't be a viable option seeing it would be extremely performance heavy.
+So for this implementation of the light mask occlusion delta value was written instead of the occlusion value.
+The occlusion delta value gets calculated by dividing the rays maximum length (distance to lightsource) by the distance to the occluder.
+This give a bit more of a smooth transition instead of the hard one and zero values.
+This will suffice for hard shadows, but does still need a lot of extra work for a soft shadow implementation.
+\pagebreak
+
+\subsubsection{Penumbra sampling}
 

### 2018-11-06 Projects/Framework/Components/TransformComponent.cpp
Worked on soft shadow implementation (currently in experimental because it has a nasty case of peter panning)

@@ -159,6 +159,7 @@ void TransformComponent::RebuildWorldMatrix()
 	glm::mat4 rotationMatrix = glm::mat4_cast(m_Rotation);
 	glm::mat4 translationMatrix = glm::translate(glm::mat4(), glm::vec3(m_Position.x, m_Position.y, m_Position.z));
 
+	//TODO: SHOULD BE MORE LIKE UNREAL
 	m_World = translationMatrix * rotationMatrix * scaleMatrix;
 
 	//Should perform scenegraph?

### 2018-11-06 Projects/Framework/Game/Application.cpp

@@ -60,8 +60,8 @@ void Application::ParseCmdLineArgs()
 
 void Application::FillPropertyManager()
 {
-	PropertyManager::GetInstance().Add("renderer_viewwidth", "1920");
-	PropertyManager::GetInstance().Add("renderer_viewheight", "1080");
+	PropertyManager::GetInstance().Add("renderer_viewwidth", "1600");
+	PropertyManager::GetInstance().Add("renderer_viewheight", "900");
 	PropertyManager::GetInstance().Add("windowtitle", "ProjectRays(OpenGL+RadeonRays)");
 }
 

### 2018-11-06 Projects/Framework/Graphics/Renderer/CL_Renderer.cpp

@@ -51,10 +51,6 @@ void CL_Renderer::OnInit(GameContext*)
 	PropertyManager::GetInstance().GetString("assetpath", m_AssetPath);
 
 	_putenv("GPU_MAX_ALLOC_PERCENT=100");
-
-	//InitCL(pGameContext);
-	//InitRadeonRays(pGameContext);
-	//InitKernels(pGameContext);
 }
 
 void CL_Renderer::OnUpdate(GameContext*)
@@ -112,6 +108,7 @@ void CL_Renderer::InitKernels(GameContext* pGameContext)
 
 	InitShadowRaysKernel(pGameContext);
 	InitLightMaskKernel(pGameContext);
+	InitSoftSampler(pGameContext);
 }
 
 void CL_Renderer::GenerateShadowRays(GameContext* pGameContext)
@@ -159,13 +156,10 @@ void CL_Renderer::GenerateLightingMask(GameContext*)
 	cl_int status = CL_SUCCESS;
 	size_t gs[] = { (size_t)m_ScreenWidth, (size_t)m_ScreenHeight };
 
-	//Acquire lightBuffer
-	status = clEnqueueAcquireGLObjects(commandQueue, 1, &m_CLGLLightBuffer, 0, nullptr, nullptr);
-	CL_ERROR_CHECK(status, "clEnqueueReleaseGLObjects failed (m_CLGLLightBuffer)");
-
 	//Get kernel from kernelprogram
-	m_LightMaskGenerator.SetArg(0, sizeof(cl_mem), &m_CLRROcclusionBuffer);
-	m_LightMaskGenerator.SetArg(1, sizeof(cl_mem), &m_CLGLLightBuffer);
+	m_LightMaskGenerator.SetArg(0, sizeof(cl_mem), &m_CLRRIntersectionBuffer);
+	m_LightMaskGenerator.SetArg(1, sizeof(cl_mem), &m_CLRRRaysBuffer);
+	m_LightMaskGenerator.SetArg(2, sizeof(cl_mem), &m_CLGLLightBuffer);
 
 	//Launch kernels on cuda cores
 	status = clEnqueueNDRangeKernel(commandQueue, m_LightMaskGenerator, 2, NULL, gs, NULL, 0, nullptr, nullptr);
@@ -173,10 +167,36 @@ void CL_Renderer::GenerateLightingMask(GameContext*)
 
 	status = clFlush(commandQueue);
[…]

### 2018-11-06 Projects/Framework/Graphics/Renderer/CL_Renderer.h

@@ -21,9 +21,11 @@ public:
 private:
 	void GenerateShadowRays(GameContext* pGameContext);
 	void GenerateLightingMask(GameContext* pGameContext);
+	void SamplePenumbra(GameContext* pGameContext);
 
 	void InitShadowRaysKernel(GameContext* pGameContext);
 	void InitLightMaskKernel(GameContext* pGameContext);
+	void InitSoftSampler(GameContext* pGameContext);
 
 	//Radeon rays context
 	RadeonRays::IntersectionApi* m_pRRContext;
@@ -36,20 +38,23 @@ private:
 	CLWProgram m_RayGenerator; //one or more kernels bundled in one
 	CLWKernel m_ShadowRayGenerator;
 	CLWKernel m_LightMaskGenerator;
+	CLWKernel m_PenumbraSampler;
 
 	//KernelData
 	cl_mem m_CLGLWorldPosBuffer;
 	cl_mem m_CLGLNormalBuffer;
 	cl_mem m_CLGLLightBuffer;
+	cl_mem m_CLGLDepthBuffer;
+	cl_mem m_CLTempLightBuffer;
 	cl_mem m_CLRRRaysBuffer;
-	cl_mem m_CLRROcclusionBuffer;
+	cl_mem m_CLRRIntersectionBuffer;
 
 	//Radeon rays data
 	RadeonRays::Buffer* m_RaysBuffer;
-	RadeonRays::Buffer* m_OcclusionBuffer;
+	RadeonRays::Buffer* m_IntersectionBuffer;
 
 	RadeonRays::ray* m_RayData;
-	int* m_OcclusionData;
+	RadeonRays::Intersection* m_IntersectionData;
 
 	int m_ScreenWidth;
 	int m_ScreenHeight;

### 2018-11-06 Projects/Framework/Graphics/Renderer/GL_Renderer.cpp

@@ -2,7 +2,7 @@
 #include "GL_Renderer.h"
 
 GL_Renderer::GL_Renderer()
-	:SingleInstance(), m_Context{}, m_pWindow{ nullptr }, m_Vsync(false), m_ScrWidth(), m_ScrHeight(), m_gBuffer(), m_WindowTitle(), m_DirectionalPos(glm::vec3(0.0f, 80.0f, 5.0f)), m_DirectionalCol(glm::vec3(1.0f, 0.9f, 0.8f))
+	:SingleInstance(), m_Context{}, m_pWindow{ nullptr }, m_Vsync(false), m_ScrWidth(), m_ScrHeight(), m_gBuffer(), m_WindowTitle(), m_DirectionalPos(glm::vec3(0.0f, 80.0f, 5.0f)), m_DirectionalCol(glm::vec3(1.0f, 0.9f, 0.8f)), m_DirectionalSize(20.0f)
 {
 }
 
@@ -141,9 +141,19 @@ unsigned int GL_Renderer::GetLightBuffer() const
 	return gLightAcc;
 }
 
+unsigned int GL_Renderer::GetDepthBuffer() const
+{
+	return gDepth;
+}
+
 const glm::vec3 & GL_Renderer::GetDirectionalLightPos() const
 {
 	return m_DirectionalPos;
+}
+
+float GL_Renderer::GetDirectionalSize() const
+{
+	return m_DirectionalSize;
 }
 
 ShaderProgram * GL_Renderer::GetLightDrawer() const
@@ -256,11 +266,10 @@ void GL_Renderer::PostInit(GameContext* pGameContext)
 	glDrawBuffers(4, attachments);
 
 	//create and attach depth buffer
-	unsigned int rboDepth;
-	glGenRenderbuffers(1, &rboDepth);
-	glBindRenderbuffer(GL_RENDERBUFFER, rboDepth);
-	glRenderbufferStorage(GL_RENDERBUFFER, GL_DEPTH_COMPONENT, m_ScrWidth, m_ScrHeight);
-	glFramebufferRenderbuffer(GL_FRAMEBUFFER, GL_DEPTH_ATTACHMENT, GL_RENDERBUFFER, rboDepth);
+	glGenRenderbuffers(1, &gDepth);
+	glBindRenderbuffer(GL_RENDERBUFFER, gDepth);
+	glRenderbufferStorage(GL_RENDERBUFFER, GL_DEPTH_COMPONENT32F, m_ScrWidth, m_ScrHeight);
[…]

### 2018-11-06 Projects/Framework/Graphics/Renderer/GL_Renderer.h

@@ -19,7 +19,10 @@ public:
 	unsigned int GetWorldPosBuffer() const;
 	unsigned int GetNormalBuffer() const;
 	unsigned int GetLightBuffer() const;
+	unsigned int GetDepthBuffer() const;
+
 	const glm::vec3& GetDirectionalLightPos() const;
+	float GetDirectionalSize() const;
 
 	ShaderProgram* GetLightDrawer() const;
 
@@ -32,7 +35,7 @@ private:
 	unsigned int quadVBO;
 
 	unsigned int m_gBuffer;
-	unsigned int gPosition, gAlbedo, gNormal, gLightAcc;
+	unsigned int gPosition, gAlbedo, gNormal, gLightAcc, gDepth;
 
 	int m_ScrWidth;
 	int m_ScrHeight;
@@ -43,6 +46,7 @@ private:
 
 	glm::vec3 m_DirectionalPos;
 	glm::vec3 m_DirectionalCol;
+	float m_DirectionalSize;
 
 	SDL_GLContext m_Context;
 	SDL_Window* m_pWindow;

### 2018-11-06 Paper/Paper_BillietStijn_RadeonRays.tex
Worked on paper (pre-final proofread)

@@
 }
-\titlepic{\includegraphics[width=\textwidth]{radeonrays.jpg}}
+\titlepic{\includegraphics[width=\textwidth]{result.png}}
 % Author
@@
 \chapter{Abstract}
-This paper was written in an effort to elaborate on hybrid solutions between modern day rasterization engines and powerful ray-tracing ones and to explain rather thoroughly how one could implement this.
-
-TODO methods
-TODO results
+This paper was written in an effort to elaborate on hybrid solutions between modern day rasterization engines and powerful ray-tracing ones and to explain rather thoroughly how one would go about implementing this.
+For the case study's implementation shadows was used to explain the main principle behind the hybrid approach.
+The hybrid rendering postpones its lighting effects through the use of deferred rendering, before going into lighting the desired lighting information is probed through the use of the raytracer.
+A stretch was made to try to improve the shadow attempt into a soft shadow one, but this attempt was feeble and did not return the wanted results.
+However the hard shadow implementation really shines seeing it performs extremely well and it even supersedes the rasterization method when it comes both speed and memory throughput.
 
@@
 \subsection{Rasterization}
-\subsubsection{OpenGL}
-\pagebreak
-\subsubsection{Loading libraries}
+\subsubsection{OpenGL Loading libraries}
 	OpenGL is a specification that has many different versions, implementations and features.
@@
 All libraries that exist out there all deliver on the same promise, meaning that the choice of library really boils down to personal preference.
+\pagebreak
 \subsubsection{GLM}
@@
 	GLM exposes a well defined interface for everything from calculus to quaternion rotations (e.g. vector operations, matrix calculations).
-\pagebreak
 \subsubsection{Window creation}
@@
 \pagebreak
-
 \subsubsection{Model loading}
@@
 \subsubsection{Raytracing using g-buffers}
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=\textwidth]{gbuffer.png}
-	\caption{Illustration of g-buffers}
-\end{figure}
 
@@
 \subsection{Deferred shading}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\textwidth]{gbuffer.png}
+\end{figure}
+
 With a standard rendering pipeline in place it was time to go ahead and turn that pipeline into a deferred one.
@@
 
-\subsubsection{Penumbra sampling}
+\subsubsection{Soft shadows}
+A naive soft shadow implementation would be to send out countless rays per light, effectively trashing performance.
+Therefore it is very tempting to try and fake that effect.
+Fortunately there is already research being done on this matter and multiple individuals\cite{kissner16} have come up with ways to fake this by using screen space sampling.
+It was only natural to go ahead and try this in our implementation.
+The first step of this algorithm was to calculate the size of the lightmask sample.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.7\textwidth]{formula.png}
+\end{figure}
+
+\begin{minipage}{0.4\textwidth}
+	This sampling was needed to try and find out how large the penumbra (shadow gradient) should be.
+Here it was a simple case of looking for the nearest non occluded (lit) pixel.
+This required a grid search to find out where in that sample that nearest pixel was.
+\end{minipage} \hfill
+\begin{minipage}{0.6\textwidth}
+	\centering
+	\includegraphics[width=0.8\textwidth]{sampler.png}
+\end{minipage}
+
+\vspace{0.5cm}
+
+\begin{minipage}{0.4\textwidth}
+	\centering
+	\includegraphics[width=0.9\textwidth]{shadowartifact.png}
+\end{minipage} \hfill
+\begin{minipage}{0.6\textwidth}
+	A large caution when using this approach is to keep in mind that this algorithm is performed in a screen space environment, meaning that the z-value does not get factored in this will bring problems when not looking straight at a surface.
+The algorithm's author\cite{kissner16} uses a highly simplified model to try and combat this, while it does alleviate some problems it's still not quite there (e.g.
+peter panning effect).\newline\par
+	The final step however was to use that distance to the nearest sample and divide it by the maximum acquirable distance within the sample area.
+\end{minipage}
+\vspace{0.5cm}
+
+This was not the result that was desired and was therefore not included in the final implementation.
 
 \chapter{Conclusion}
+When it came to rasterization setting up a deferred pipeline was easy, there was quite a lot of groundwork that was done already.
+Rasterization in its essence is already quite performant so no additional optimization were needed.
+The raytracing part of this implementation came with some challenges regarding not losing too much performance while still trying to gain from raytracing.
+The median road that was taken here was to make sure there was a zero copy protocol between the different api's to ensure as little time as possible was wasted towards buss contention.
+Fortunately for us radeon rays already came with quite a big optimization being the sorting of the meshes in BVH's, so for this implementation no major additional optimization were needed.\newline\par
+
+A hybrid pipeline is definitely the way to go seeing the performance impact of raytracing and the limited amount of dedicated hardware acceleration (e.g.
+No Nvidia RTX cards).
+The main upsides of a hybrid pipeline being that the large cost of the primary rays can totally be augmented and that the renderer can now use world information rather than only surface information.
+This on its own can cause a ray traced effect to be even more effective than a rasterization one (e.g.
+shadows).
+The only real downside of a full hybrid approach is the limited amount of scalability seeing that this almost forces the renderer to always use ray tracing for lighting effects, this can be alleviated by implementing a highly modular pipeline, but this would require a lot of work to get right.
 
@@
 	\url{https://www.scratchapixel.com/lessons/3d-basic-rendering/ray-tracing-generating-camera-rays}
+
+	\bibitem{kissner16}
+	Michael Kissner,
+	2016,
+	Ray traced soft shadows in real time in Spellwrath,
+	\url{https://www.imgtec.com/blog/ray-traced-soft-shadows-in-real-time-spellwrath/}
 \end{thebibliography}

### 2019-05-27 Paper/Paper_BillietStijn_RadeonRays.tex
Improved hard shadow implementation in awaition for soft shadow changes

@@
 \begin{figure}[h]
-	\caption{Illustration of different ray-casting solutions}
+	\caption{Illustration of different ray-casting solutions \cite{keepmaxing16}}
 	\centering
@@
 \includegraphics[width=\textwidth]{rasterization.jpg}
-\caption{Rasterization graphics pipeline.}
+\caption{Rasterization graphics pipeline. \cite{learnopengl18}}
 \end{figure}
@@
 
-\subsubsection{Graphics pipeline}
+\subsubsection{Graphics pipeline}\cite{pipeline18}
 With rasterization objects on screen are created from meshes.
@@
 	\includegraphics[width=0.8\textwidth]{coordinate_systems.png}
-	\caption{Model space to screen space.}
+	\caption{Model space to screen space. \cite{learnopengl18}}
 	\label{fig:coordinates}
@@
 	\includegraphics[width=\textwidth]{primitives.png}
-	\caption{List of possible OpenGL primitives.}
+	\caption{List of possible OpenGL primitives. \cite{ratcliffe05}}
 \end{figure}
@@
 	\includegraphics[width=0.7\textwidth]{geometryshader.jpg}
-	\caption{Purpose of geometry shader.}
+	\caption{Purpose of geometry shader. \cite{rastergrid00}}
 \end{figure}
@@
 	\includegraphics[width=\textwidth]{culling.png}
+	\cite{mayer18}
 \end{minipage} \hfill
@@
 	\includegraphics[width=\textwidth]{frustrum.png}
+	\cite{frustrumyourdictionary18}
 \end{minipage} \hfill
@@
 	\centering
-	\includegraphics[width=0.85\textwidth]{Zbuffer.png}
+	\includegraphics[width=0.85\textwidth]{Zbuffer.png}\cite{Zbuffering18}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{alpha.png}
+	\includegraphics[width=0.8\textwidth]{alpha.png}\cite{yourdictionary18}
 \end{figure}
@@
 \subsection{Raytracing}
-\subsubsection{OpenCL}
+\subsubsection{OpenCL}\cite{lapere16}
 Open Computing Language is a framework for writing programs that execute across compute devices.
@@
 Under this section you'll find a quick rundown of aforementioned execution 
-model.
+model.\cite{lapere16}
 \pagebreak
@@
 	\includegraphics[width=0.7\textwidth]{memorymodel.png}
-	\caption{Illustration of the memory execution model.}
+	\caption{Illustration of the memory execution model. \cite{lapere16}}
 \end{figure}
@@
 	\includegraphics[width=\textwidth]{sharegroups.png}
-	\caption{OpenCL/OpenGL interop context}
+	\caption{OpenCL/OpenGL interop context \cite{apple18}}
 \end{figure}
@@
 		\includegraphics[width=0.9\textwidth]{assimpstructure.png}
+		\caption{Illustation of AssImp model structure. \cite{learnopengl18}}
 	\end{figure}
@@
 	\includegraphics[width=\textwidth]{fvsdef.png}
-	\caption{Difference between forward and deferred pipeline.}
+	\caption{Difference between forward and deferred pipeline. \cite{owens18}}
 \end{figure}
@@
 	\includegraphics[width=\textwidth]{fvsdefgraph.png}
-	\caption{Lighting performance comparison between forward and deferred.}
+	\caption{Lighting performance comparison between forward and deferred. \cite{vanoosten18}}
 \end{figure}
@@
 	\includegraphics[width=0.6\textwidth]{ndc.png}
-	\caption{Normalized device coordinates for our triangle}
+	\caption{Normalized device coordinates for our triangle. \cite{learnopengl18}}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=\textwidth]{vertexattributepointer.png}
+	\includegraphics[width=\textwidth]{vertexattributepointer.png}\cite{learnopengl18}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=0.7\textwidth]{formula.png}
+	\includegraphics[width=0.7\textwidth]{formula.png}\cite{kissner16}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{sampler.png}
+	\includegraphics[width=0.8\textwidth]{sampler.png}\cite{kissner16}
 \end{minipage}
@@
 
+	\bibitem{keepmaxing16}
+	Keep Maxing,
+	2016,
+	Mastering Vray Lighting \& Materials - 052 Understanding Vray Render Optimization,
+	\url{https://www.youtube.com/watch?v=A3z9v2MVRvw}
+
+	\bibitem{learnopengl18}
+	Learn OpenGL,
+	2018,
+	Learn OpenGL,
+	\url{https://learnopengl.com/}
+
+	\bibitem{pipeline18}
+	Wikipedia,
+	2018,
+	Graphics pipeline,
+	\url{https://en.wikipedia.org/wiki/Graphics_pipeline}
+
+	\bibitem{yourdictionary18}
+	yourdictionary
+	2018,
+	alpha-blending,
+	\url{http://www.yourdictionary.com/alpha-blending}
+
+	\bibitem{Zbuffering18}
+	WikiPedia
+	2018,
+	Z-buffering,
+	\url{https://en.wikipedia.org/wiki/Z-buffering}
+
+	\bibitem{ratcliffe05}
+	Stephen Ratcliffe
+	2005,
+	Review of OpenGL,
+	\url{http://vision-simulator.tripod.com/report/section_7.htm}
+
+	\bibitem{rastergrid00}
+	RasterGrid
+	2000,
+	history-of-hardware-tessellation,
+	\url{http://rastergrid.com/blog/2010/09/history-of-hardware-tessellation/}
+
+	\bibitem{mayer18}
+	A. Julian Mayer,
+	2018,
+	Dynamic Occlusion Culling,
+	\url{https://pdfs.semanticscholar.org/5cab/0fac256b81b71c8d85a51881f7daf52ab7f3.pdf}
+
+	\bibitem{frustrumyourdictionary18}
+	yourdictionary
+	2018,
+	view frustrum,
+	\url{http://www.yourdictionary.com/viewing-frustum}
+
+	\bibitem{lapere16}
+	Sam Lapera,
+	2016,
+	OpenCL path tracing tutorial 1: Firing up OpenCL
+	\url{http://raytracey.blogspot.com/2016/11/opencl-path-tracing-tutorial-1-firing.html}
+
+	\bibitem{apple18}
+	Apple,
+	2018,
+	Sharing Data Between OpenCL and OpenGL
+	\url{https://developer.apple.com/library/archive/documentation/Performance/Conceptual/OpenCL_MacProgGuide/shareGroups/shareGroups.html}
+
+	\bibitem{owens18}
+	Ben Owens,
+	2018,
+	Forward rendering vs deferred
+	\url{https://gamedevelopment.tutsplus.com/articles/forward-rendering-vs-deferred-rendering--gamedev-12342}
+
+	\bibitem{vanoosten18}
+	Jeremiah van Oosten,
+	2018,
+	Forward vs Deferred vs Forward+ Rendering with DirectX 11
+	\url{https://www.3dgep.com/forward-plus/}
+
 	\bibitem{voica14}
@@
 
-	\bibitem{pipeline18}
-	Wikipedia,
-	2018,
-	Graphics pipeline,
-	\url{https://en.wikipedia.org/wiki/Graphics_pipeline}
-
 	\bibitem{scratchapixel18}

### 2019-05-27 Projects/Framework/Components/CameraComponent.cpp

@@ -2,7 +2,7 @@
 #include "CameraComponent.h"
 
 CameraComponent::CameraComponent()
-	:m_NearPlane(0.1f), m_FarPlane(100.0f), m_FOV(60.0f), m_UsePerspective(true), m_Orthosize(1.0f)
+	:m_NearPlane(0.1f), m_FarPlane(200.0f), m_FOV(60.0f), m_UsePerspective(true), m_Orthosize(1.0f)
 {
 }
 

### 2019-05-27 Projects/Framework/Game/Application.cpp

@@ -60,8 +60,8 @@ void Application::ParseCmdLineArgs()
 
 void Application::FillPropertyManager()
 {
-	PropertyManager::GetInstance().Add("renderer_viewwidth", "1920");
-	PropertyManager::GetInstance().Add("renderer_viewheight", "1080");
+	PropertyManager::GetInstance().Add("renderer_viewwidth", "1280");
+	PropertyManager::GetInstance().Add("renderer_viewheight", "720");
 	PropertyManager::GetInstance().Add("windowtitle", "ProjectRays(OpenGL+RadeonRays)");
 }
 
@@ -99,11 +99,14 @@ void Application::Render()
 	auto pCLRenderer = m_pGameContext->m_pCLRenderer;
 	auto pGameMode = m_pGameContext->m_pGamemode;
 
+	//Deferred render context
 	pGLRenderer->Begin();
-	pGameMode->Draw(m_pGameContext);
-	pCLRenderer->RaytracedShadows(m_pGameContext);
-	pGLRenderer->LightPass();
-	pGLRenderer->End();
+	{
+		pGameMode->Draw(m_pGameContext);
+		pCLRenderer->RaytracedShadows(m_pGameContext);
+		pGLRenderer->LightPass();
+	}
+	pGLRenderer->End(m_pGameContext);
 }
 
 void Application::OnUpdate(GameContext* pGameContext)

### 2019-05-27 Projects/Framework/Graphics/Model/Model.cpp

@@ -111,9 +111,13 @@ Mesh Model::ProcessMesh(aiMesh* pMesh)
 	auto ambientMaps = LoadMaterialTextures(material, aiTextureType_AMBIENT, TextureType::AMBIENT);
 	textures.insert(textures.end(), ambientMaps.begin(), ambientMaps.end());
 
-	// 3. ambient maps
+	// 3. normal maps
 	auto normalMaps = LoadMaterialTextures(material, aiTextureType_NORMALS, TextureType::NORMAL);
 	textures.insert(textures.end(), normalMaps.begin(), normalMaps.end());
+
+	// 4. specular maps
+	auto specularMaps = LoadMaterialTextures(material, aiTextureType_SPECULAR, TextureType::SPECULAR);
+	textures.insert(textures.end(), specularMaps.begin(), specularMaps.end());
 
 	return Mesh(vertices, indices, textures);
 }

### 2019-05-27 Projects/Framework/Graphics/Renderer/CL_Renderer.cpp

@@ -35,6 +35,8 @@ using namespace RadeonRays;
 
 CL_Renderer::CL_Renderer()
 	:SingleInstance<CL_Renderer>()
+    ,m_AngularExtent(2.0f)
+    ,m_TileSize(4)
 {
 }
 
@@ -51,10 +53,6 @@ void CL_Renderer::OnInit(GameContext*)
 	PropertyManager::GetInstance().GetString("assetpath", m_AssetPath);
 
 	_putenv("GPU_MAX_ALLOC_PERCENT=100");
-
-	//InitCL(pGameContext);
-	//InitRadeonRays(pGameContext);
-	//InitKernels(pGameContext);
 }
 
 void CL_Renderer::OnUpdate(GameContext*)
@@ -120,7 +118,7 @@ void CL_Renderer::GenerateShadowRays(GameContext* pGameContext)
 	cl_int status = CL_SUCCESS;
 	auto dirLightPos = pGameContext->m_pGLRenderer->GetDirectionalLightPos();
 	cl_float4 light_cl = { dirLightPos.x, dirLightPos.y, dirLightPos.z, 1.0f };
-	size_t gs[] = { (size_t)m_ScreenWidth, (size_t)m_ScreenHeight };
+    size_t gs[] = { (size_t)m_ScreenWidth, (size_t)m_ScreenHeight };
 
 	//Acquire WorldPosBuffer
 	status = clEnqueueAcquireGLObjects(commandQueue, 1, &m_CLGLWorldPosBuffer, 0, nullptr, nullptr);
@@ -131,10 +129,12 @@ void CL_Renderer::GenerateShadowRays(GameContext* pGameContext)
 	CL_ERROR_CHECK(status, "clEnqueueAcquireGLObjects failed (m_CLGLNormalBuffer)");
 
 	//Send new information to shader
-	m_ShadowRayGenerator.SetArg(0, sizeof(cl_float4), &light_cl);
-	m_ShadowRayGenerator.SetArg(1, sizeof(cl_mem), &m_CLRRRaysBuffer);
-	m_ShadowRayGenerator.SetArg(2, sizeof(cl_mem), &m_CLGLWorldPosBuffer);
-	m_ShadowRayGenerator.SetArg(3, sizeof(cl_mem), &m_CLGLNormalBuffer);
+    m_ShadowRayGenerator.SetArg(0, sizeof(cl_float4), &light_cl);
[…]

### 2019-05-27 Projects/Framework/Graphics/Renderer/CL_Renderer.h

@@ -18,6 +18,9 @@ public:
 
 	void RaytracedShadows(GameContext* pGameContext);
 	RadeonRays::IntersectionApi* GetRaysAPI() const;
+
+    float* GetAngularExtent() { return &m_AngularExtent; };
+    unsigned int* GetTileSize() { return &m_TileSize; };
 private:
 	void GenerateShadowRays(GameContext* pGameContext);
 	void GenerateLightingMask(GameContext* pGameContext);
@@ -54,4 +57,7 @@ private:
 	int m_ScreenWidth;
 	int m_ScreenHeight;
 	std::string m_AssetPath;
+
+    float m_AngularExtent;
+    unsigned int m_TileSize;
 };

### 2019-05-27 Projects/Framework/Graphics/Renderer/GL_Renderer.cpp

@@ -81,20 +81,20 @@ void GL_Renderer::LightPass()
 	glUniform3fv(directionalColVar, 1, &m_DirectionalCol[0]);
 }
 
-void GL_Renderer::End()
+void GL_Renderer::End(GameContext* pGameContext)
 {
-	// finally render quad
+	//finally render quad
 	RenderQuad();
 
-	// copy content of geometry's depth buffer to default framebuffer's depth buffer
-	// -----------------------------------------------------------------------------
+	//copy content of geometry's depth buffer to default framebuffer's depth buffer
+	//-----------------------------------------------------------------------------
 	glBindFramebuffer(GL_READ_FRAMEBUFFER, m_gBuffer);
 	glBindFramebuffer(GL_DRAW_FRAMEBUFFER, 0);
 
-	// write to default framebuffer
-	// blit to default framebuffer. Note that this may or may not work as the internal formats of both the FBO and default framebuffer have to match.
-	// the internal formats are implementation defined. This works on all of my systems, but if it doesn't on yours you'll likely have to write to the 		
-	// depth buffer in another shader stage (or somehow see to match the default framebuffer's internal format with the FBO's internal format).
+	//write to default framebuffer
+	//blit to default framebuffer. Note that this may or may not work as the internal formats of both the FBO and default framebuffer have to match.
+	//the internal formats are implementation defined. This works on all of my systems, but if it doesn't on yours you'll likely have to write to the 		
+	//depth buffer in another shader stage (or somehow see to match the default framebuffer's internal format with the FBO's internal format).
 	glBlitFramebuffer(0, 0, m_ScrWidth, m_ScrHeight, 0, 0, m_ScrWidth, m_ScrHeight, GL_DEPTH_BUFFER_BIT, GL_NEAREST);
 	glBindFramebuffer(GL_FRAMEBUFFER, 0);
 
@@ -103,7 +103,7 @@ void GL_Renderer::End()
 	ImGui_ImplSDL2_NewFrame(m_pWindow);
 	ImGui::NewFrame();
 
-	ImGuiOnDraw();
+	ImGuiOnDraw(pGameContext);
 
 	ImGui::Render();
[…]

### 2019-05-27 Projects/Framework/Graphics/Renderer/GL_Renderer.h

@@ -10,7 +10,7 @@ public:
 
 	void Begin();
 	void LightPass();
-	void End();
+	void End(GameContext* pGameContext);
 
 	const SDL_GLContext& GetGLContext() const;
 	SDL_Window* GetWindow() const;
@@ -19,6 +19,7 @@ public:
 	unsigned int GetWorldPosBuffer() const;
 	unsigned int GetNormalBuffer() const;
 	unsigned int GetLightBuffer() const;
+	void SetDirectionalLightPos(const glm::vec3& newPos);
 	const glm::vec3& GetDirectionalLightPos() const;
 
 	ShaderProgram* GetLightDrawer() const;
@@ -27,7 +28,7 @@ public:
 	virtual void PostInit(GameContext* pGameContext) override;
 private:
 	void RenderQuad();
-	void ImGuiOnDraw();
+	void ImGuiOnDraw(GameContext* pGameContext);
 	unsigned int quadVAO = 0;
 	unsigned int quadVBO;
 

### 2019-05-27 Projects/Game/Gamemode/GMSandbox.cpp

@@ -8,6 +8,9 @@
 
 using Super = BaseGamemode;
 
+bool GMSandbox::_RunLightCycle = false;
+bool GMSandbox::_AutoPanCamera = false;
+
 GMSandbox::GMSandbox()
 	:BaseGamemode()
 {
@@ -28,13 +31,13 @@ void GMSandbox::PreInit(GameContext* pContext)
 	Super::OnInit(pContext);
 
 	//Level
-	auto pRenderTest = new RenderTestScene();
+	//auto pRenderTest = new RenderTestScene();
 	auto pSponzaScene = new SponzaScene();
-	auto pShadowTestScene = new ShadowTestScene();
+	//auto pShadowTestScene = new ShadowTestScene();
 
-	pContext->m_pSceneManager->Add(pRenderTest);
+	//pContext->m_pSceneManager->Add(pRenderTest);
 	pContext->m_pSceneManager->Add(pSponzaScene);
-	pContext->m_pSceneManager->Add(pShadowTestScene);
+	//pContext->m_pSceneManager->Add(pShadowTestScene);
 
 	pContext->m_pSceneManager->SetActive("SponzaScene");
 
@@ -60,4 +63,10 @@ void GMSandbox::PreInit(GameContext* pContext)
 	auto pLook = new LookAround(pContext);
 	auto output = EventFactory::CreateDesc(SDL_KEYDOWN, SDLK_p);
 	pContext->m_pInputManager->AddInputAction(output, pLook);
+
+	auto pLightCycleToggle = new ToggleLightCycle();
+	pContext->m_pInputManager->AddInputAction(EventFactory::CreateDesc(SDL_KEYUP, SDLK_k), pLightCycleToggle);
+
+	auto pAutoPanCamera = new AutoPanCamera();
+	pContext->m_pInputManager->AddInputAction(EventFactory::CreateDesc(SDL_KEYUP, SDLK_l), pAutoPanCamera);
 }

### 2019-05-27 Projects/Game/Gamemode/GMSandbox.h

@@ -16,5 +16,7 @@ public:
 	void OnUpdate(GameContext* pContext) override;
 	void PreInit(GameContext* pContext) override;
 
+	static bool _RunLightCycle;
+	static bool _AutoPanCamera;
 private:
 };

### 2019-05-27 Projects/Game/Gamemode/InputActions.cpp

@@ -1,5 +1,6 @@
 #include "GamePCH.h"
 #include "InputActions.h"
+#include "GMSandbox.h"
 
 Forward::Forward(GameContext* pContext)
 	:m_pGameContext(pContext)
@@ -110,4 +111,24 @@ void LookAround::execute()
 
 	glm::quat rotQuat = qYaw * qPitch;
 	m_pTransformComp->Rotate(rotQuat);
+}
+
+ToggleLightCycle::ToggleLightCycle()
+{
+}
+
+void ToggleLightCycle::execute()
+{
+	Logger::GetInstance().LogInfo("Light cycle started \n");
+	GMSandbox::_RunLightCycle = !GMSandbox::_RunLightCycle;
+}
+
+AutoPanCamera::AutoPanCamera()
+{
+}
+
+void AutoPanCamera::execute()
+{
+	Logger::GetInstance().LogInfo("Camera auto panning on \n");
+	GMSandbox::_AutoPanCamera = !GMSandbox::_AutoPanCamera;
 }

### 2019-05-27 Projects/Game/Gamemode/InputActions.h

@@ -79,3 +79,21 @@ private:
 	TransformComponent* m_pTransformComp;
 	glm::vec2 m_PrevMousePos;
 };
+
+class ToggleLightCycle : public Command
+{
+public:
+	ToggleLightCycle();
+
+	virtual void execute() override;
+private:
+};
+
+class AutoPanCamera : public Command
+{
+public:
+	AutoPanCamera();
+
+	virtual void execute() override;
+private:
+};

### 2019-05-27 Projects/Game/Scenes/RenderTestScene.cpp

@@ -35,17 +35,12 @@ void RenderTestScene::PreInit(GameContext* pContext)
 	//Load model and fetch meshes
 	m_pModel = new Model("Models/Suicune/Suicune.obj");
 	auto pSuicuneMesh = &(m_pModel->GetMeshes()[0]);
-	auto pPlatformMesh = &(m_pModel->GetMeshes()[1]);
 
 	std::vector<Texture> textureVect{};
 	textureVect.push_back(Texture("Textures/Suicune/SuicuneSaix_A.tga.png", TextureType::DIFFUSE));
 	textureVect.push_back(Texture("Textures/Suicune/SuicuneSaix_N.tga.png", TextureType::NORMAL));
-	textureVect.push_back(Texture("Textures/Suicune/SuicuneSaix_AO.tga.png", TextureType::AMBIENT));
+	textureVect.push_back(Texture("Textures/Suicune/SuicuneSaix_M.tga.png", TextureType::SPECULAR));
 	pSuicuneMesh->SetTextures(textureVect);
-
-	std::vector<Texture> otherTextureVect{};
-	otherTextureVect.push_back(Texture("Textures/spnza_bricks_a_diff.png", TextureType::DIFFUSE));
-	pPlatformMesh->SetTextures(otherTextureVect);
 
 	//Load shaders and bundle in shaderprogram
 	auto pShaderProgram = new ShaderProgram("Shaders/Deferred_Geometry.vs", "Shaders/Deferred_Geometry.fs");
@@ -56,20 +51,11 @@ void RenderTestScene::PreInit(GameContext* pContext)
 
 	auto pRayShapeComponent = new RayShapeComponent(pSuicuneMesh);
 	m_pTestObject->AddComponent(pRayShapeComponent);
-
-	//Translate object and add to scene
-	auto pPlatformObject = new GameObject();
-	pMeshDrawComp = new MeshDrawComponent(pPlatformMesh, pShaderProgram);
-	pPlatformObject->AddComponent(pMeshDrawComp);
-
-	pRayShapeComponent = new RayShapeComponent(pPlatformMesh);
-	pPlatformObject->AddComponent(pRayShapeComponent);
-	m_pTestObject->Add(pPlatformObject);
 }
 
 void RenderTestScene::PostInit(GameContext * pContext)
 {
 	Super::PostInit(pContext);
 
[…]

### 2019-05-27 Projects/Game/Scenes/SponzaScene.cpp

@@ -1,10 +1,11 @@
 #include "GamePCH.h"
 #include "SponzaScene.h"
+#include "Gamemode/GMSandbox.h"
 
 using Super = Scene;
 
 SponzaScene::SponzaScene()
-	:Scene("SponzaScene")
+	:Scene("SponzaScene"), m_SunAngle()
 {
 }
 
@@ -40,5 +41,35 @@ void SponzaScene::PreInit(GameContext * pContext)
 void SponzaScene::PostInit(GameContext * pContext)
 {
 	Super::PostInit(pContext);
-	pContext->m_pActiveCamera->GetGameObject()->GetTransform()->Translate(0.0f, 0.0f, 0.0f);
+	pContext->m_pActiveCamera->GetGameObject()->GetTransform()->Translate(-55.0f, 15.0f, 0.0f);
+	pContext->m_pActiveCamera->GetGameObject()->GetTransform()->Rotate(0.0f, -90.0f, 0.0f, false);
+}
+
+void SponzaScene::OnUpdate(GameContext* pContext)
+{
+	Super::OnUpdate(pContext);
+
+	if (GMSandbox::_AutoPanCamera)
+	{
+		auto pTransform = pContext->m_pActiveCamera->GetGameObject()->GetTransform();
+		auto prevPostion = pTransform->GetWorldPosition();
+		pTransform->Translate(prevPostion + glm::vec3(0.5f, 0.0f, 0.0f) * pContext->elapsedSec);
+	}
+
+	if (GMSandbox::_RunLightCycle)
+	{
+		m_SunAngle += 5.0f * pContext->elapsedSec;
+		if (m_SunAngle > 360.0f) m_SunAngle = 0;
+
+		//x = r * cos(delta_angle)
+		//y = r * sin(delta_angle)
+
+		auto glRenderer = pContext->m_pGLRenderer;
+		auto prevLightPos = glRenderer->GetDirectionalLightPos();
+
+		float newXCoord = 20.0f * cos(glm::radians(m_SunAngle));
+		float newZCoord = 20.0f * sin(glm::radians(m_SunAngle));
+		auto newLightPos = glm::vec3(newXCoord, prevLightPos.y, newZCoord);
+		glRenderer->SetDirectionalLightPos(newLightPos);
+	}
 }

### 2019-05-27 Projects/Game/Scenes/SponzaScene.h

@@ -9,6 +9,8 @@ public:
 private:
 	virtual void PreInit(GameContext* pContext) override;
 	virtual void PostInit(GameContext* pContext) override;
+	virtual void OnUpdate(GameContext* pContext) override;
 
 	std::vector<Mesh> m_Meshes;
+	float m_SunAngle;
 };
